### 2025-05-16 paper/summary-family-access-measures-table.tex

{\tiny
\begin{longtable}{|p{2.5cm}|p{2.5cm}|p{2.5cm}|p{3cm}|p{3cm}|}
\hline
\textbf{Name of Case and Variant} & \textbf{Constraint Explanation and Balancing Factor} & \textbf{Proportional Allocation Factor} & \textbf{Measure Equation} & \textbf{Interpretation} \\
\hline
\endfirsthead

\hline
\textbf{Name of Case and Variant} & \textbf{Constraint Explanation and Balancing Factor} & \textbf{Proportional Allocation Factor} & \textbf{Measure Equation} & \textbf{Interpretation} \\
\hline
\endhead

Unconstrained Accessible Opportunities or Population ($V_i^0$, $M_j^0$)
& No constraints; marginals not preserved.
& None
& $V_i^0 = \sum_j O_j f(c_{ij})$;

$M_j^0 = \sum_i P_i f(c_{ij})$

& Impedance-weighted values (e.g., "opportunities × decay"); no marginal control. \\
\hline

Total Constrained Accessible Opportunities or Population ($V_i^T$, $M_j^T$)
& Single balancing factor preserves total marginal.

$K^T = \frac{D}{\sum_i V_i^0}$;

$\hat{K}^T = \frac{O}{\sum_j M_j^0}$

& $\kappa_i^T = \sum_j K^T f(c_{ij})$; $\hat{\kappa}_j^T = \sum_i \hat K^T f(c_{ij})$
& $V^T_i = \kappa_i^T \sum_j W^{(2)}_j$;

$M_j^T = \hat{\kappa}_j^T \sum_i W^{(1)}_i$

& Values reflect total regional opportunities or population. \\
\hline

Singly Constrained Accessible Opportunities ($V_i^S$)
& Destination-side constraint:

$B_j = \frac{1}{\sum_i W_i^{(1)} f(c_{ij})}$
& $\kappa^S_i = W_i^{(1)} B_j$
& $V^S_i = \kappa^S_i \sum_j D_j$
& Shares of destination-side opportunities based on origin population and impedance. \\
\hline

Singly Constrained Accessible Population ($M_j^S$)
& Origin-side constraint:

$A_i = \frac{1}{\sum_j W_j^{(2)} f(c_{ij})}$

& $\hat \kappa^S_j = W_j^{(2)} A_i$
& $M_j^S = \hat \kappa^S_j \sum_i P_i$
& Shares of origin population based on destination mass and impedance. \\
\hline

Doubly Constrained Access ($V_{ij}^D$ or $M_{ij}^D$)
& Both marginals constrained via $A_i$, $B_j$ simultaneously.
& —
& $V_{ij}^D = A_i B_j P_i O_j f(c_{ij})$
& Realized interactions between population and opportunities (i.e., access). \\
\hline

\end{longtable}
}