### 2025-08-01 paper_plos/paper_plos_V0.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%

\PassOptionsToPackage{table}{xcolor}

\documentclass[
  10pt,
  letterpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% marvosym package for additional characters
\usepackage{marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% Using natbib instead
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}
\usepackage{float}
\floatplacement{table}{H}
\floatplacement{figure}{H}
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage[numbers,square,comma]{natbib}
\bibliographystyle{plos2015}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A family of accessibility measures derived from spatial interaction principles},
  pdfauthor={Anastasia Soukhov; Rafael H. M. Pereira; Christopher D. Higgins; Antonio Páez},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{A family of accessibility measures derived from spatial
interaction
principles} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
Anastasia Soukhov\textsuperscript{1}, Rafael H. M.
Pereira\textsuperscript{2}, Christopher D.
Higgins\textsuperscript{3*}, Antonio Páez\textsuperscript{1}
\\
\bigskip
\textbf{1} McMaster University, School of Earth, Environment \&
Society, Hamilton, Canada, \\ \textbf{2} Institute for Applied Economic
Research - Ipea, Data Science Division, Brasília,
Brazil, \\ \textbf{3} University of Toronto Scarborough, Department of
Human Geography, Toronto, Canada, 
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
\dag Deceased

% Group/Consortium Author Note
\textpilcrow Membership list can be found in the Acknowledgments
sections

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* cd.higgins@utoronto.ca

\end{flushleft}

\section*{Abstract}
Transportation planning has long prioritized the efficiency of movement,
or mobility. However, the concept of accessibility represents a more
comprehensive evolution, shifting focus from mere movement to the
potential to reach (i.e., spatially interact) with desired destinations.
Despite growing recognition of accessibility-based planning approaches,
the concept remains fragmented, with inconsistent definitions and
unclear interpretations. This work's aim is to clarify and unify the
concept of accessibility by connecting it into spatial interaction
modeling. We demonstrate that widely used mobility and accessibility
models, such as gravity-based accessibility and spatial interaction
models, share common theoretical roots. From this foundation, this paper
offers three contributions: (A) we introduce a family of accessibility
measures within the principles of spatial interaction, and (B) formally
define four members of the family, namely the `unconstrained' measure
(i.e., Hansen-type accessibility), the `total constrained' measure
(i.e., a constrained version of the Hansen-type accessibility), the
`singly constrained' measure (i.e., related to the popular two step
floating catchment approach - 2SFCA), and the `doubly constrained'
measure representing realized interactions or `access', effectively
equal to the doubly constrained spatial interaction model; and (C) we
demonstrate the interpretability advantages of the family, as these
constrained accessibility measures yield values in units of the number
of potential ``opportunities for spatial interaction'' or ``population
for spatial interaction'' for each zone and zonal flow. The family of
accessibility measures proposed here clarifies the concept of
`potential' in accessibility, demonstrates theoretical and formulaic
linkages across popular accessibility and spatial interaction models,
and reintroduces measurement units into accessibility measures. By doing
so, we believe this family of measures can unlock a clearer, more
interpretable, and cohesive foundation for accessibility analysis.

\linenumbers

\section{Introduction}\label{introduction}

In the early nineteenth century, industrializing cities grew rapidly: so
did traffic congestion and the development of transportation planning
practice focused primarily on mobility. In this newly founded practice,
access to destinations was treated as a by-product of movement.
Following the rise of the automobile and significant investments in
transportation infrastructure after World War II, this mobility-oriented
approach has led to some problematic outcomes. Specifically, the car
became seen as the ultimate mobility tool, helping to foster the
development of low-density, single-use residential neighborhoods and
entrenching an automobility mono-culture within transportation systems
\citep{miller_collaborative_2011, lavery_driving_2013}. Decades of
planning for this automobility mono-culture, still often characterized
by road and highway expansion, have been marked by increased travel
costs and environmental burdens, with limited impact on enhancing the
ease with which people can reach destinations
\citep{farber_running_2011, handyACCESSIBILITYVSMOBILITYENHANCING2002, paez_healthcare_2010}.
In response, transportation researchers have increasingly advocated for
the adoption of accessibility as a planning criterion, in contrast to
traditional mobility-oriented transportation planning approaches which
translate into indicators that benchmark movement (e.g., vehicle
kilometres traveled, intersection through traffic, etc.) which are not
necessarily linked to improved accessibility
\citep{silvaAccessibilityInstrumentsPlanning2017, paez_developing_2013, handy2020, elgeneidyMakingAccessibilityWork2022}.

Accessibility may be defined as the ``potential of opportunities for
{[}spatial{]} interaction'' \citep{hansen1959}, in contrast to mobility
which is simply, movement in itself or `spatial interaction'. Mobility
has been the basis of traditional transportation planning approaches
\citep{ortuzar_2011_modelling}. While accessibility brings a more
holistic understanding of combined transportation and land use systems,
incorporating both the explicit consideration of destination as
opportunity and the concept of \emph{potential}
\citep{handyMeasuringAccessibilityExploration1997}.

The growing interest in accessibility has been accompanied by a boom in
scholarly research using different methods and focusing on different
research contexts; it has grown to include studies of access to
employment
\citep[e.g.,][]{karstEvaluationAccessibilityImpacts2003, grengs2010job, paez_jobs_2013, merlin2017competition, tao_investigating_2020},
health care
\citep[e.g.,][]{luo2003, paez_healthcare_2010, wan2012three, delamater2013spatial, boisjoly2017informality, pereira_2021_geographic},
green spaces
\citep{reyesAccessibility2014, rojas_accessibility_2016, liang_novel_2024},
schools
\citep[e.g.,][]{williams_disparities_2014, romanillosAccessibilitySchoolsSpatial2018, marques_accessibility_2021},
social contacts
\citep[e.g.,][]{neutens_human_2007, farberActivitySpacesMeasurement2012, farber_2013_social},
and regional economic analysis
\citep[e.g.,][]{vickermanAccessibility1999, lopezMeasuring2008, ribeiro_road_2010, gutierrez_evaluating_2011}
among many other domains of application. In other words, accessibility
analysis is used today to broadly understand the potential to reach (or
spatially interact with) opportunities that are important to people
\citep{ferreiraReenactingMobilityAccessibility2020}. However, despite
its growth in popularity in scholarly works, challenges remain with
respect to the more widespread adoption of accessibility in planning
practice. For instance, the diversity of accessibility definitions has
been flagged by van Wee \citep{vanweeAccessible2016}, Handy
\citep{handy2020}, and Kapatsila et al.
\citep{kapatsila_resolving_2023}. Further, difficulties in the
interpretability and communicability of outputs has also been noticed by
many authores, including Geurs and van Wee
\citep{geursAccessibilityEvaluationLanduse2004}, van Wee
\citep{vanweeAccessible2016}, and Ferreira and Papa
\citep{ferreiraReenactingMobilityAccessibility2020}.

The adoption of accessibility in planning practice is not necessarily
made easier when potential adopters have to contend with a plethora of
definitions, each seemingly more sophisticated but less intuitive than
the last \citep{kapatsila_resolving_2023}. At a high level, Geurs and
van Wee \citep{geursAccessibilityEvaluationLanduse2004} identify four
families of accessibility measures: infrastructure-, place-, person-,
and utility-based. Of the place-based family, which is the focus of this
paper, the menu has grown to include gravity-based accessibility
\citep[e.g.,][]{hansen1959, pirie_measuring_1979}, cumulative
opportunities
\citep[e.g.,][]{wachs_physical_1973, pirie_measuring_1979, ye_spatial_2018},
modified gravity \citep[e.g.,][]{SchuurmanMeasuring2010}, 2-Step
Floating Catchment Areas \citep[e.g.,][]{luo2003}, Enhanced 2-Step
Floating Catchment Areas \citep[e.g.,][]{luoEnhanced2009}, 3-Stage
Floating Catchment Areas \citep[e.g.,][]{wan2012three}, Modified 2-Step
Floating Catchment Areas \citep[e.g.,][]{delamater2013spatial}, inverse
2-Step Floating Catchment Areas \citep[e.g.,][]{wang_2sfca_2021}, and
n-steps Floating Catchment Areas \citep{liang_novel_2024}. How is a
practitioner to choose among this myriad options? What differences in
accessibility scores should matter, and how should they be communicated?
\citep[see][ p.~14]{vanweeAccessible2016}.

Here, we seek to address the breath of place-based accessibility's
definitions and lack of interpretability by demonstrating that
mobility-oriented models--such as the commonly used gravity-based
accessibility \citep[e.g.,][]{hansen1959} and the spatial interaction
model \citep[e.g.,][]{wilson1971}-- are rooted in the same spatial
interaction modeling foundation. In fact, we propose that accessibility
can be specified using spatial interaction principles as a \emph{family
of measures}, akin to the family of spatial interaction models
introduced in Wilson \citep{wilson1971} that can be defined using
balancing factors that constrains values based on known information
about the land-use.

This work aims to offer three contributions: (1) it introduces a family
of accessibility measures within the principles of spatial interaction;
(2) it formally defines three \emph{constrained} accessibility measures
by reintroducing Wilson-analogous balancing factors. These measures are
the total constrained accessibility measure, the singly constrained
accessibility measure, and the doubly constrained measure, which are
explicitly connected to popular measures such as the Hansen-type
accessibility \citep{hansen1959}, the popular competition approach of
the 2-Step Floating Catchment Area (2SFCA)
\citep{shenLocationCharacteristicsInnercity1998, luo2003}, and the
concept of market potential
\citep{harris_market_1954, vickermanAccessibilityAttractionPotential1974}.
These Wilson-analogous balancing factors introduced are also defined in
order of increasing restrictiveness to shed light on the role of
\emph{potential} in accessibility and access. (3) This work also
demonstrates that the introduction of balancing factors makes
accessibility measures easier to interpret and to communicate by
restoring the measurement units to the resulting raw accessibility
values. Each zonal and zonal flow value from a constrained accessibility
measure is always in interpretable units, namely the number of
`opportunities for spatial interaction' or `population for spatial
interaction'. This is in contrast to conventional accessibility
measures, particularly gravity based measures, that yield values in
units of `opportunities weighted by some representation of travel
friction'.

To achieve these stated objectives, this paper contends that
accessibility research must reconnect with its spatial interaction
origins. Particularly, we argue that an important aspect of spatial
interaction modelling--namely, constraining the results to match
empirical observations--was never effectively reincorporated into
accessibility analysis. Empirical constraints were embraced by early
spatial interaction literature following the work of Wilson
\citep{wilson1971}, but this stream of literature tended to flow
separately from research inspired by Hansen's \citep{hansen1959}
accessibility. The application of Wilson's \citep{wilson1971} empirical
constraints supported the development of various spatial interaction
models that remain relevant in research and practice today
\citep{ortuzar_2011_modelling}. However, the same cannot be said of the
contemporary accessibility literature, where empirical constraints were
not explicitly adopted. We argue that the absence of empirical
constraints (and their attendant proportionality constants) has
contributed to some of accessibility analysis' interpretability issues;
for instance, the fuzziness of insights beyond simple proportional
statements like `higher-than' or `lower-than'
\citep{millerAccessibilityMeasurementApplication2018}. Moreover, without
constraints we lose track of what accessibility measures, that is, the
number of opportunities that can be spatially interacted with. Without a
clear sense of measurement units, comparability between accessibility
measures, across cities, and transport modes, may also become
compromised.

Fortunately, accessibility and spatial interaction modelling literature
share common headwaters, and the latter has given careful attention to
measurement units and their interpretability. It is by looking to the
past that we believe accessibility analysis can newly wade into the
future. We continue this work in the following section by tracing the
development of accessibility from its origins in spatial interaction:
from the Newtonian gravitational expression in Eavenstein
\citep{ravensteinLawsMigration1889} through to the seminal accessibility
work of Hansen \citep{hansen1959}. We then present evidence for a
narrative highlighting the marked divergence between accessibility and
spatial interaction modelling research after the work of Wilson
\citep{wilson1971}. Next, we hark back to Wilson's \citep{wilson1971}
spatial interaction models, and use them to derive a family of
accessibility measures based on analogous constraints. We illustrate
members of this family of constrained accessibility measures with a
simple numerical example. We then conclude by discussing the uses of
these measure and their interpretation.

\section{Newtonian's roots of human spatial interaction
research}\label{newtonians-roots-of-human-spatial-interaction-research}

The patterns of people's movement in space have been a subject of
scientific inquiry for at least a century and a half. From as far back
as Henry C. Carey's \emph{Principles of Social Science}
\citep{careyPrinciplesSocialScience1858}, a concern with the scientific
study of human spatial interaction can be observed. It was in this work
where Carey stated that ``man {[}is{]} the molecule of society {[}and
their interaction is subject to{]} the direct ratio of the mass and the
inverse one of distance'' \citep[pp.~37-38]{mckeanManual1883}. This
statement shows how investigations into human spatial interaction have
often been explicitly coloured by the features of Newton's Law of
Universal Gravitation, first posited in 1687's \emph{Principia
Mathematica} and expressed as in Eq~\ref{eq-phys-grav-prop}.

\begin{equation}\phantomsection\label{eq-phys-grav-prop}{
F_{ij} \propto \frac{M_i M_j} {D_{ij}^{2}}
}\end{equation}

To be certain, the expression above is one of proportionality, and is
also the most famous in all of science. It states that the force of
attraction \(F\) between a pair of bodies \(i\) and \(j\) is directly
\emph{proportional} to the product of their masses \(M_i\) and \(M_j\),
and inversely \emph{proportional} to the square of the distance between
them \(D_{ij}\). Direct proportionality means that as the product of the
masses increases, so does the force. Likewise, inverse proportionality
means that as the distance increases, the force decreases.
Eq~\ref{eq-phys-grav-prop}, however, does not quantify the magnitude of
the force. To do so, an empirical constant, a.k.a. the gravitational
constant, is required to convert the proportionality into an equality,
ensuring that values of the force \(F\) in Eq~\ref{eq-phys-grav-prop}
match the observed force of attraction between masses. In other words,
Eq~\ref{eq-phys-grav-prop} needs to be \emph{constrained} using
empirical data. Ultimately, the equation for the force is as seen in
Eq~\ref{eq-phys-grav}, where \(G\) is an empirically calibrated
proportionality constant:

\begin{equation}\phantomsection\label{eq-phys-grav}{
F_{ij} = G \frac{M_i M_j} {D_{ij}^{2}}
}\end{equation}

Newton's initial estimate of \(G\) was based on a speculation that the
mean density of earth was between five or six times that of water, an
assumption that received support after Hutton's experiments of 1778
\citep[p.~783]{hutton_xxxiii_1778}. Still, it took over a century from
the publication of \emph{Principia} to refine the estimate of the
proportionality constant to within 1\% accuracy, with Cavendish's 1798
experiment \citep{cavendish_xxi_1798}.

\section{Early research on human spatial interaction: from Ravenstein
(1889) to Stewart
(1948)}\label{early-research-on-human-spatial-interaction-from-ravenstein-1889-to-stewart-1948}

Since the 1880s to the 1940s, a number of researchers theoretically and
empirically attempted to characterise human spatial interaction as some
force of attraction \(F\) that is directly proportion to the masses
\(M_i\) and \(M_j\) and inversely proportional by their separation
distance. This concept was captured with different expressions, but all
tie back to the same Newtonian gravity analogy, although not all of them
included a proportionality constant in their formulation.

Following Carey's \emph{Principles} of 1858, research into human spatial
interaction continued in different contexts. In the late 1880s,
Ravenstein proposed some ``Laws of Migration'' based on his empirical
analysis of migration flows in various countries
\citep{ravensteinLawsMigration1885, ravensteinLawsMigration1889}. In
these works, Ravenstein posited 1) a directly proportional relationship
between migration flows and the size of destinations (i.e., centres of
commerce and industry), and 2) an inversely proportional relationship
between the size of flows and the separation between origins and
destinations. As with Carey, these propositions echo Newton's
gravitational laws. Over time, other researchers discovered similar
relationships. For example, Reilly \citep{reilly1929methods} formulated
a law of retail gravitation, expressed in terms of equal attraction to
competing retail destinations that could be understood as `potential
trade territories'. Later, Zipf proposed a \(\frac{P_1P_2}{D}\)
hypothesis for the case of information
\citep{zipfDeterminantsCirculationInformation1946}, intercity personal
movement \citep{zipfHypothesisIntercityMovement1946}, and goods movement
by railways \citep{zipfHypothesisCaseRailway1946}. The
\(\frac{P_1P_2}{D}\) hypothesis stated that the magnitude of flows was
proportional to the product of the populations of settlements, and
inversely proportional to the distance between them.

A common feature of these early investigations of human spatial
interaction is that a proportionality constant similar to \(G\) in
Eq~\ref{eq-phys-grav} was never considered. Of the researchers cited
above, only Reilly and Zipf expressed their hypotheses in mathematical
terms. Reilly's hypothesis was presented in the following form:

\begin{equation}\phantomsection\label{eq-reilly}{
B_a = \frac{(P_a P_T)^N}{D_{aT}^n}
}\end{equation}

\noindent where \(B_a\) is the amount of business drawn to \(a\) from
\(T\), \(P_a\) and \(P_T\) are the populations of the two settlements,
and \(D_{aT}\) is the distance between them. Quantity \(N\) was chosen
by Reilly in a somewhat \emph{ad hoc} fashion as 1, and he used
empirical observations of shoppers to choose a value of \(n = 2\).

Zipf, on the other hand, wrote his hypothesis in mathematical form as:

\begin{equation}\phantomsection\label{eq-zipf}{
C^2 = \frac{P_1 P_2}{D_{12}}
}\end{equation}

\noindent where \(C\) is the volume of goods exchanged between \(1\) and
\(2\), \(P_1\) and \(P_2\) are the populations of the two settlements,
and \(D_{12}\) is the distance between them.

After Carey, it is in Stewart's work on the principles of demographic
gravitation that we find the strongest connection yet to Newton's law
\citep{stewartDemographicGravitationEvidence1948}. This may relate to
academic backgrounds; where Stewart was a physicist while Ravenstein,
Reilly, and Zipf were social scientists. Besides awareness of preceding
research (he cites both Reilly and Zipf as predecessors in the analysis
of human spatial interaction), Stewart appears to have been the first
author to express his theorized relationships for human spatial
interaction with a proportionality constant \(G\), as follows:

\begin{equation}\phantomsection\label{eq-stewart-force}{
F = G\frac{(\mu_1N_1)(\mu_2N_2)}{d_{12}^2} = G\frac{M_1 M_2}{d_{12}^2} 
}\end{equation}

\noindent Where:

\begin{itemize}
\tightlist
\item
  \(F\) is the \emph{demographic force}
\item
  \(N_1\) and \(N_2\) are the numbers of people of in groups 1 and 2
\item
  \(\mu_1\) and \(\mu_2\) are so-called \emph{molecular weights}, the
  attractive weight of groups 1 and 2
\item
  \(M_1 = \mu_1N_1\) and \(M_2 = \mu_2N_2\) are the demographic masses
  at 1 and 2
\item
  \(d_{12}^2\) is the distance between \(1\) and \(2\)
\item
  And finally \(G\), a constant that Stewart ``left for future
  determination''
  \citep[p.~34]{stewartDemographicGravitationEvidence1948}
\end{itemize}

In addition to demographic force, Stewart defined a measure of the
``population potential'' of group \(2\) with respect to group \(1\). In
other words, the potential number of people from location \(2\) that
could visit location \(1\), as follows:

\begin{equation}\phantomsection\label{eq-stewart-population-potential}{
V_1 = G\frac{M_2}{d_{12}}
}\end{equation}

For a system with more than two population bodies, Stewart formulated
the population potential at \(i\) as follows (after arbitrarily assuming
that \(G=1\)):

\begin{equation}\phantomsection\label{eq-stewart-population-potential-integral}{
V_i = \int\frac{D}{r} ds
}\end{equation}

\noindent where \(D\) is the population density over an infinitesimal
area \(ds\) and \(r\) is the distance to \(i\). In
Eq~\ref{eq-stewart-population-potential-integral}, \(D ds\) gives an
infinitesimal count of the population, say \(dm\), and so, after
discretizing space, Eq~\ref{eq-stewart-population-potential-integral}
can be rewritten as:

\begin{equation}\phantomsection\label{eq-stewart-population-potential-sum}{
V_i = \sum_j M_jd_{ij}^{-1}
}\end{equation}

Alerted readers will notice that
Eq~\ref{eq-stewart-population-potential-sum}, with some re-organization
of terms, is formally equivalent to our modern definition of
accessibility popularized by Hansen \citep{hansen1959} in the late
1950s.

Stewart's formulation of demographic force, developed in the context of
what he called ``social physics'' \citep{stewartPrinciples1947}, was
problematic. It had issues with inconsistent mathematical notation. More
seriously though, Stewart's work was permeated by a view of humans as
particles following physical laws, but tinted by unscientific and racist
ideas. For instance, he assumed that the molecular weight \(\mu\) of the
average American was one, but ``presumably\ldots much less than
one\ldots.for an Australian aborigine'' {[}p.~35{]}. Stewart's ideas
about ``social physics'' soon fell out of favour among social
scientists, but not before influencing the nascent field of
accessibility research, as detailed next.

\section{Hansen's gravity-based accessibility to
today}\label{grav-to-today}

From Stewart \citep{stewartDemographicGravitationEvidence1948}, we
arrive to 1959 and Walter G. Hansen, whose work proved to be
exceptionally influential in the accessibility literature
\citep{hansen1959}. In this seminal paper, Hansen defined accessibility
as ``the potential of opportunities for interaction\ldots{} a
generalization of the population-over-distance relationship or
\emph{population potential} concept developed by Stewart
\citep{stewartDemographicGravitationEvidence1948}'' (p.~73). As well as
being a student of city and regional planning at the Massachusetts
Institute of Technology, Hansen was also an engineer with the Bureau of
Roads, and preoccupied with the power of transportation to shape land
uses in a very practical sense. Hansen \citep{hansen1959} focused on
Stewart's \citep{stewartDemographicGravitationEvidence1948}
\emph{population potential} (expressed in
Eq~\ref{eq-stewart-population-potential-sum}), and not on the other
formulaic contributions and objectionable aspects of ``social physics''.
Hansen \citep{hansen1959} recast Stewart's population potential to
reflect accessibility, a model of human behaviour useful to capture
regularities in mobility patterns. Hansen \citep{hansen1959} replaced
\(M_j\) in Eq~\ref{eq-stewart-population-potential-sum} with
\emph{opportunities} to derive an \emph{opportunity potential}, or more
accurately, a \emph{potential of opportunities for interaction} as
follows:

\begin{equation}\phantomsection\label{eq-accessibility}{
S_{i} = \sum_j \frac{O_j }{d_{ij}^\beta}
}\end{equation}

A contemporary rewriting of Eq~\ref{eq-accessibility} accounts for a
variety of impedance functions beyond the inverse power \(d^{-\beta}\):

\begin{equation}\phantomsection\label{eq-accessibility-general}{
S_{i} = \sum_j O_j  f(d_{ij})
}\end{equation}

\(S_{i}\) in Eq~\ref{eq-accessibility} is a measure of the accessibility
from site \(i\). This is a function of \(O_j\) (the mass of
opportunities at \(j\)), \(d_{ij}\) (the cost, e.g., distance or travel
time, incurred to reach \(j\) from \(i\)), and \(\beta\) (a parameter
that modulates the friction of cost). Today, Hansen is frequently cited
as the father of modern accessibility analysis
\citep[e.g.,][]{reggianiGuestEditorialNew2011}, and Hansen-type
accessibility is commonly referred to as the gravity-based accessibility
measure.

However, Hansen's use of Stewart's \emph{population potential} measure
included one crucial omission that afflicts the literature to this day.
The omission is that between Stewart
\citep{stewartDemographicGravitationEvidence1948} and Hansen
\citep{hansen1959}, the proportionality constant \(G\) in
Eq~\ref{eq-stewart-population-potential} vanished. This constant was not
explicitly addressed in Hansen \citep{hansen1959} and accessibility
research continues to evolve without it. Since Hansen
\citep{hansen1959}, accessibility analysis has been widely used in
numerous disciplines but, to our knowledge, the proportionality constant
has remained forgotten, with no notable developments to explicitly
acknowledge or determine it.

The omission of this constant generates a fundamental problem for the
measurement unit of accessibility estimates, which undermines the
interpretation, communication and comparability of accessibility
analysis. Those reading Hansen \citep{hansen1959} must recall that
Stewart \citep{stewartDemographicGravitationEvidence1948} had set the
proportionality constant \(G\) to 1, with a note that ``\(G\) {[}was{]}
left for future determination: a suitable choice of other units can
reduce it to unity'' {[}p.~34{]}. In practice, the persistent omission
of the constant in accessibility analyses means that \(G\) continues to
be implicitly set to \(1\), even when the fundamental relationship in
accessibility is proportionality (e.g.,
\(S_{i} \propto \sum_j g(O_j)f(d_{ij})\)) and not equality {[}for
instance, see the formula for accessibility at the top of Figure 1 in Wu
and Levinson \citep{wuUnifyingAccess2020}{]}. The direct consequence is
that without a proportionality constant, the units of \(S_i\) remain
unclear: the unit of ``potential of opportunity for interaction'' is
left free to change as \(\beta\) is calibrated. For example, if
\(c_{ij}\) is distance in meters, it will be number of opportunities per
\(m^{\beta}\) when \(f(c_{ij}) = d^{-\beta}\) but number of
opportunities per \(e^{-\beta m}\) when \(f(c_{ij}) = e^{-\beta m}\).
This undermines the comparability of accessibility metrics with
different decay functions, and renders their results difficult to
understand and communicate. The Hansen-style accessibility estimates
found in the literature, therefore, are better thought of as ordinal
measures of potential that can only be interpreted in terms of higher
and lower accessibility, but which has not palpable meaning
\citep{millerAccessibilityMeasurementApplication2018}.

\section{Wilson's family of spatial interaction
models}\label{wilsons-family-of-spatial-interaction-models}

On the other side of the Atlantic, Alan G. Wilson was developing
related, yet parallel work. In his groundbreaking study
\citep{wilson1971}, Wilson defined a general spatial interaction model
as follows. While accessibility was characterised as an associated
concept of `potential', the primary focus was on modelling observed
spatial interaction:
\begin{equation}\phantomsection\label{eq-phys-gravity-model}{
T_{ij} = k W_i^{(1)} W_j^{(2)} f(c_{ij})
}\end{equation}

The model in Eq~\ref{eq-phys-gravity-model} posits a quantity \(T_{ij}\)
that represents a value in a matrix of flows of size \(n \times m\),
that is, between \(i = 1,s, n\) origins and \(j = 1,s, m\) destinations.
The quantities \(W_i^{(1)}\) and \(W_j^{(2)}\) are proxies for the
masses at \(i=1,s,n\) origins and \(j=1,s,m\) destinations. The
super-indices \((1)\) and \((2)\) are meant to indicate that these
masses could be different things, i.e., \(W_i^{(1)}\) could be
populations, and \(W_j^{(2)}\) hectares of park space. Finally,
\(f(c_{ij})\) is some function of travel cost \(c_{ij}\) which reflects
travel impedance. In this way, \(T_{ij}\) explicitly measures
\emph{interaction} in the unit of trips, and the role of \(k\) is to
ensure that the system-wide sum of \(T_{ij}\) represents the total flows
in the data. In other words, \(k\) is a scale parameter that makes the
overall amount of flows identical to the magnitude of the phenomenon
being modeled. In other words, it balances the units, in a conceptually
similar sense as the gravitational constant in Newton's Law of Universal
Gravitation.

Traditionally, the development of the spatial interaction model put an
emphasis on the interpretability of the results
\citep{kirbyNormalizingFactorsGravity1970, wilsonSTATISTICALTHEORYSPATIAL1967, wilson1971}.
But instead of relying on the heuristic of Newtonian gravity (e.g., some
interaction between a mass at \(i\) and a mass at \(j\) separated by
some distance), Wilson's approach was to maximise the entropy of the
system. Entropy maximisation in this case achieves stable results as a
statistical average that represents the population. The approach works
by assuming undifferentiated individual interactions, and assessing
their probabilities of making a particular journey. The result of
Eq~\ref{eq-phys-gravity-model} then is a statistical average
\citep{wilson1971, seniorGravityModellingEntropy1979}.

To ensure that \(T_{ij}\) in Eq~\ref{eq-phys-gravity-model} is in the
unit of trips (our unit of origin-destination spatial interaction),
additional knowledge about the system is required. At the very least,
this framework assumes that the total number of trips in the system
\(T\) is known, and therefore:

\begin{equation}\phantomsection\label{eq-constraint0-gravitymodel}{
\sum_i\sum_j T_{ij} = T
}\end{equation}

Additional information can be introduced. For example, when information
is available about the total number of trips produced by each origin,
\(W_i^{(1)}\) is represented as \(O_i\) and the following constraint can
be used:

\begin{equation}\phantomsection\label{eq-constraint1-gravitymodel}{
\sum_j T_{ij} = O_i
}\end{equation}

Alternatively, if there is information available about the total number
of trips attracted by each destination, \(W_j^{(2)}\) is represented as
\(D_j\) and the following constraint can be used:

\begin{equation}\phantomsection\label{eq-constraint2-gravitymodel}{
\sum_i T_{ij} = D_j
}\end{equation}

It is also possible to have information about both \(O_i\) and \(D_j\),
in which case both constraints could be imposed on the model at once.

Using information about the system that satisfy these constraints fully,
partially or not at all, a family of spatial interaction models can be
derived based on Eq~\ref{eq-phys-gravity-model}. \(K\) is specified
depending on the applied constraint(s). In the framework introduced in
Wilson \citep{wilson1971}, three constrained versions are specified: the
first being a case where to results only match the total volume of
interaction, the second being a singly constrained case, and the third a
doubly constrained case.

In the first, Eq~\ref{eq-constraint1-gravitymodel} and
Eq~\ref{eq-constraint2-gravitymodel} do not hold. In practical terms,
this means that the total number of trips predicted by the model must be
equal to sum of all flows from origins i to destinations j. The
balancing constant \(K\) in this case is
\citep[see][]{cliff_evaluating_1974, fotheringham_spatial_1984}:

\begin{equation}\phantomsection\label{eq-total-flow-balancing-factor}{
K=\frac{T}{\sum_i\sum_j T_{ij}}
}\end{equation}

In the second case, only one of Eq~\ref{eq-constraint1-gravitymodel} or
Eq~\ref{eq-constraint2-gravitymodel} hold. The resulting models are, in
Wilson's terms, singly constrained. When only
Eq~\ref{eq-constraint1-gravitymodel} holds, entropy maximisation leads
to the following production-constrained model:

\begin{equation}\phantomsection\label{eq-production-constrained-gravitymodel}{
T_{ij} = A_i O_i W_j^{(2)} f(c_{ij})
}\end{equation}

Notice how, in this model, the proxy for the mass at the origin
\(W_i^{(1)}\) is replaced with \(O_i\), representing what we know about
the system, the spatial interaction outbound flow, i.e., outbound trips
produced at \(i\). Also, there is no longer a single system-wide
proportionality constant, but rather a set of proportionality constants
(i.e., balancing factors) specific to origins. For this model, the
balancing factors ensure that Eq~\ref{eq-constraint1-gravitymodel} is
satisfied, meaning that the sum of predicted flows from one origin going
to all destinations must equal the known mass at that origin \(O_i\)
i.e., the total number of outbound trips. Satisfying this constraint
also implicitly fulfills the total constraint
(Eq~\ref{eq-constraint0-gravitymodel}), since the sum of \(O_i\) values
across all origins equals the total number of trips. This model is
useful when trips ends are unknown but the number of trips originating
from each location is known and the total of these trips represents all
trips in the system. The balancing factors for the
production-constrained model are solved for each origin \(A_i\), and
according to Wilson are:

\begin{equation}\phantomsection\label{eq-production-constrained-balancing-factor}{
A_i = \frac{1}{\sum_j W_j^{(2)} f(c_{ij})}
}\end{equation}

The attraction-constrained model is similar to the
production-constrained model as it is also singly constrained but from
the perspective of the mass at the destination. From the
attraction-constrained model, the proxy for the mass at the destination
\(W_j^{(2)}\) is now replaced with \(D_j\), representing the spatial
interaction inbound flow, i.e., trips attracted at the destination and
takes the following form:

\begin{equation}\phantomsection\label{eq-attraction-constrained-gravitymodel}{
T_{ij} = B_j D_j W_i^{(1)} f(c_{ij})
}\end{equation}

In this model (Eq~\ref{eq-attraction-constrained-gravitymodel}), the
balancing factors ensure that Eq~\ref{eq-constraint2-gravitymodel} is
satisfied (hence the total constraint
Eq~\ref{eq-constraint0-gravitymodel} is as well), meaning that the sum
of predicted flows going to one destination from all origins must equal
the known mass of that destination \(D_j\) i.e., the total number of
inbound trips to \(j\). This should hold for all destinations. As
before, destination-specific proportionality constants (i.e., balancing
factors) \(B_j\) were derived by Wilson as:

\begin{equation}\phantomsection\label{eq-attraction-constrained-balancing-factor}{
B_j = \frac{1}{\sum_i W_i^{(1)} f(c_{ij})}
}\end{equation}

The third case in the family of spatial interaction models is the
production-attraction constrained model. In this case, both
Eq~\ref{eq-constraint1-gravitymodel} and
Eq~\ref{eq-constraint2-gravitymodel} hold simultaneously. These
constraints ensure that the sum of predicted flows from one origin to
all destination, and the predicted flows going to one destination from
all origins must equal the known mass of the origin \(O_i\) and of the
destination \(D_j\). This should hold for all origins and destinations.
The resulting model is, in Wilson's terms, doubly constrained, and takes
the following form:

\begin{equation}\phantomsection\label{eq-doubly-constrained-gravitymodel}{
T_{ij} = A_i B_j O_i D_j f(c_{ij})
}\end{equation}

In this model, both proxies for the masses are replaced with the known
masses, that is, the trips produced by the origin and the trips
attracted by the destination. There are now two sets of mutually
dependent proportionality constants:

\begin{equation}\phantomsection\label{eq-doubly-constrained-balancing-factors}{
\begin{array}{l}
A_i = \frac{1}{\sum_j B_j D_j f(c_{ij})}\\
B_j = \frac{1}{\sum_i A_i O_i f(c_{ij})}
\end{array}
}\end{equation}

Derivation of these models is demonstrated in detail elsewhere
\citep[e.g.,][]{ortuzar_2011_modelling, wilsonSTATISTICALTHEORYSPATIAL1967}.
It is worth noting, however, that although Wilson's approach is built on
a different conceptual foundation than the old reference to Newtonian
gravity, the work succeeded at identifying the steps from
proportionality to equality to yield variations of proportionality
constants, including the one that eluded Stewart
\citep{stewartDemographicGravitationEvidence1948} and that has been
overlooked in almost all subsequent accessibility research. Why was this
key element of spatial interaction models potentially ignored in
accessibility research? In the next section we aim to address this
question.

\section{Accessibility and spatial interaction modelling: two divergent
research
streams}\label{accessibility-and-spatial-interaction-modelling-two-divergent-research-streams}

The work of Hansen \citep{hansen1959} and Wilson \citep{wilson1971}
responded to important developments, in particular a need ``to meet the
dictates and needs of public policy for strategic land use and
transportation planning'' \citep{battyChronicleScientificPlanning1994}.
These dictates and needs were far from trivial. In the United States
alone, the Federal-Aid Highway Act of 1956 authorized the creation of
the U.S. Interstate Highway System, with a budget that ultimately
exceeded one hundred billion dollars (equivalent to over \$600 billion
in 2023)
\citep{weinerUrbanTransportationPlanning2016, mdotMnDOTJoins2007}.
Spatial interaction modelling was incorporated into institutional
modelling practices meant to ``predict and provide'', i.e., predict
travel demand and supply transportation infrastructure
\citep{kovatch1971modeling, weinerUrbanTransportationPlanning2016}.
Accessibility, at the time, did not quite have that power, as it did not
quantify trips, but rather something somewhat more elusive: it predicts
the less tangible ``potential'' for spatial interaction with
opportunities. In this way, where spatial interaction modelling became a
key element of transportation planning practice, accessibility remained
a somewhat more academic pursuit, and the two streams of literature only
rarely connected.

\begin{figure}[H]

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{paper_plos_files/figure-pdf/fig-docs-per-year-1.pdf}

}

\caption{\label{fig-docs-per-year}Historical pattern of publication:
documents per year.}

\end{figure}%

To illustrate this point, we conducted a bibliographic analysis of the
literature that cites Hansen \citep{hansen1959}, Wilson
\citep{wilson1971}, or both. We retrieved all relevant documents using
the Web of Science ``Cited References'' functionality, and the digital
object identifiers of Hansen \citep{hansen1959} and Wilson
\citep{wilson1971}. As a result of this search, we identified 1,788
documents that cite Hansen \citep{hansen1959}, only 258 documents that
cite Wilson \citep{wilson1971}, and 76 that cite both. The earliest
document in this corpus dates to 1976 and the most recent is from 2025.
The number of documents per year appears in Fig~\ref{fig-docs-per-year},
where we see the frequency of documents over a span of almost fifty
years. In particular, we notice the remarkable growth in the number of
papers that cite Hansen \citep{hansen1959} compared to those that cite
Wilson \citep{wilson1971} since the year 2000.

As noted, literature that cites both Wilson \citep{wilson1971} and
Hansen \citep{hansen1959} are sparse (only 3.6\% of the corpus,
visualised in pink in Fig~\ref{fig-docs-per-year}). After reading the
works, we can also discern that they too are divergent, with one stream
focused on developing accessibility (i.e., \emph{potential for} spatial
interaction) and another on spatial interaction. The focuses of these
divergent streams contribute this paper's broader hypothesis that the
concepts of accessibility and spatial interaction have remained largely
disconnected, and at times, improperly conflated.

On one hand, the stream of literature focused on spatial interaction
models inspired by Wilson \citep{wilson1971} and which cite both Hansen
\citep{hansen1959} and Wilson \citep{wilson1971}, tends to contribute to
understanding how accessibility is interpreted and incorporated in
spatial interaction models. These works treat separate spatial
interaction and accessibility as a seperate but related phenomenon.
Specifically, some early works interpret the spatial interaction model's
balancing factors (Eq~\ref{eq-production-constrained-balancing-factor}
or Eq~\ref{eq-doubly-constrained-balancing-factors}) as the inverse of
Hansen's \citep{hansen1959} model
\citep{harrisEquilibriumValuesDynamics1978, leonardiOptimumFacilityLocation1978, fotheringhamSPATIALSTRUCTUREDISTANCE1981, fotheringhamSpatialCompetitionAgglomeration1985},
recognizing it as a ``common sense'' approach
\citep[p.~99]{morrisAccessibilityIndicatorsTransport1979} to including
accessibility in the spatial interaction model, though further
exploration of its relationship is warranted
\citep{battyMethodResiduesUrban1976}. Some authors have explored this
relationship, for instance as in Fotheringham
\citep{fotheringhamSpatialCompetitionAgglomeration1985} who demonstrates
how the spatial interaction model may insufficiently explain spatial
patterns, and suggests that explicitly defining destinations'
accessibility as a variable within the model may remedy the issue (e.g.,
the \emph{competition destination} model). Other works used both Hansen
\citep{hansen1959} and Wilson's \citep{wilson1971} framework in
conjunction, such as in defining location-allocation problems in
operations research
\citep{leonardiOptimumFacilityLocation1978, beaumontLocationallocationProblemsPlane1981},
estimating trip flows (or some other spatial interaction flows)
alongside accessibility
\citep[e.g.,][]{clarke2002deriving, grengs2004measuring, turk2019socio},
or considering accessibility within spatial interaction models, in line
with Fotheringham's
\citep{fotheringhamSpatialCompetitionAgglomeration1985} demonstration
\citep[e.g.,][]{beckers2022incorporating}. Other works departed from
Hansen's \citep{hansen1959} definition and aligned with spatial
interaction in different ways, such as using micro-economic consumer
behaviour concepts to express potential for spatial interaction
\citep{morrisAccessibilityIndicatorsTransport1979, leonardiRandomUtilityDemand1984}.

On the other hand, there is another subset of literature that cite both
Hansen \citep{hansen1959} and Wilson \citep{wilson1971} that is
accessibility-focused. We categorise their citation of Wilson
\citep{wilson1971} for three general reasons. Firstly, a group of these
works cite Wilson \citep{wilson1971} as attribution for using
context-dependent travel cost functions
\citep[e.g.,][]{weibullNumericalMeasurementAccessibility1980, handyMeasuringAccessibilityExploration1997, kwan1998space, shenLocationCharacteristicsInnercity1998, ashiru2003space, rau2012spatial, pan2013impacts, margaridacondecomelhoradoImpactMeasuringInternal2016, caschili2015accessibility, grengs2015nonwork, pan2020measuring, chia2020extending, roblot2021participation, sharifiasl2023incorporating, kharel2024examining}.
These works do not necessarily comment on spatial interaction
explicitly. Secondly, another group of this accessibility-focused
``both'' citing literature \emph{does} associate spatial interaction as
defined in Wilson \citep{wilson1971} with accessibility's potential for
spatial interaction more explicitly
\citep[e.g.,][]{millerMeasuringSpaceTimeAccessibility1999, giuliano2010accessibility, grengs2010intermetropolitan, grengs2010job, grengs2012equity, levine2012does, levinson2012positive, tong2015transportation, liu2015spatial, he2017measuring, wuUnifyingAccess2020, ng2022reflection, naqavi2023mobility, suel2024measuring}.
We agree accessibility and spatial interaction are related topics:
accessibility is an expression of its \emph{potential} and the Wilson
\citep{wilson1971} paper briefly touches on the concept. However, in
some of this literature, Hansen \citep{hansen1959} and Wilson
\citep{wilson1971} are co-cited as both being `gravity models'
\citep[e.g.,][]{liu2004accessibility, dai2017visualization, shen2019segregation, chia2020extending},
perhaps revealing the murkiness of the distinction between spatial
interaction and the \emph{potential for} spatial interaction in the
literature. Thirdly, there is a group of accessibility-focused works
that interpret Hansen's \citep{hansen1959} model as the singly- or
doubly- constrained spatial interaction model's inverse balancing factor
\citep[e.g.,][]{vickermanAccessibilityAttractionPotential1974}. This
group cities the earlier spatial interaction works that make this
connection and is especially prominent in the investigation of
competitive accessibility topics
\citep[e.g.,][]{karstEvaluationAccessibilityImpacts2003, geurs2006accessibility, willigers2007accessibility, el2011place, curtis2010planning, manaugh2012makes, chen2013regional, alonso2014labour, albacete2017measuring, sahebgharani2019computing, mayaud2019future, allenMeasureCompetitiveAccess2020, levinsonGeneralTheoryAccess2020, marwal2022literature, su2023untangling}.
As outlined in preceding sections, we argue interpreting the singly- or
doubly- constrained spatial interaction model's balancing factor as
accessibility yields output values that are similarly plagued by
interpretability issues.

Lastly, as an extension of the third reason within this group, only the
works of Soukhov et al.
\citep{soukhovIntroducingSpatialAvailability2023, soukhovMultimodalSpatialAvailability2024}
use Wilson's \citep{wilson1971} balancing factors as a method for
maintaining constraints on opportunities within the context of
competitive accessibility. These works introduce the balancing factors
as a mechanisms to ensure that opportunities at each destination are
proportionally allocated to each zone (based on the proportion of
population seeking opportunities and the relative travel impedance).
This is to ensure that each zonal accessibility value is the sum of this
proportional allocation from each destination, and that all zonal values
ultimately sum to the total number of opportunities in the region.
However, these balancing factors were deduced intuitively. These works
do not explicitly state that the mathematical formulation of the
equations are effectively equivalent to Wilson's singly constrained
model (derived from entropy maximization). This equivalence is only
discovered in hindsight, as will be demonstrated in this study. These
two works also do not discuss other constrained cases that will be
addressed in the next section.

\section{A family of accessibility
measures}\label{a-family-of-accessibility-measures}

In this section, we introduce a family of accessibility measures
inspired by Wilson's spatial interaction modelling framework. At the end
of this section, Table 1 presents a summary, where each member of the
family of accessibility measure is explained in plain language,
alongside their balancing factor, proportional allocation factor and
mathematical equation and value interpretations.

As argued in the preceding section, the streams of research on
accessibility and spatial interaction modelling have evolved as largely
separate streams with little contact since Hansen \citep{hansen1959} and
Wilson \citep{wilson1971}. This may explain why the constraints and
associated balancing factors of spatial interaction models did not cross
over to accessibility analysis. This is intriguing since Wilson made an
effort to connect developments in spatial interaction modelling to
accessibility, noting for instance, the denominator of the
proportionality constants specific to origins
(Eq~\ref{eq-production-constrained-balancing-factor}) is the inverse of
balancing factor \(A_i\) \citep[ p.~10]{wilson1971}:

\begin{equation}\phantomsection\label{eq-Ai-as-accessibility}{
S_i = \frac{1}{A_i} = \sum_j W_j^{(2)} f(c_{ij})
}\end{equation}

Understanding \(S_i\) as the inverse of Wilson's \(A_i\) does not
uncover any new meaning for \(S_i\) itself. Indeed, mathematically it is
true, \(A_i\)'s role in Wilson's general model
\(T_{ij} = k W_i^{(1)} W_j^{(2)} f(c_{ij})\) is that of a balancing
factor \(k\) i.e., keeping units balanced and proportionality based on
constraints. Understanding \(S_i\) itself as a balancing factor is not
wholly helpful as Hansen and (Stewart before him) defined accessibility
as a partial sum of the demographic force \(F\) or the system-wide
population potential of opportunities for spatial interaction (i.e.,
Eq~\ref{eq-stewart-population-potential-sum} representing
\(V_i = \sum_j \frac{M_j}{d_{ij}}\) after losing \(G\)).

Hence, we propose stepping back to introduce a revised definition of
accessibility: the \emph{constrained} potential for spatial interaction.
Once we bring back Wilson's proportionality constant \(k\) into the
picture, we can define the \emph{potential for spatial interaction}
between two locations \(i\) and \(j\) is as follows:

\begin{equation}\phantomsection\label{eq-access-01}{
V_{ij} = k W_j^{(2)} f(c_{ij})
}\end{equation}

\noindent where \(V_{ij}\) is the potential for interaction from \(i\)
to \(j\). The accessibility from origin \(i\) can then be summarised as
as a partial sum of the potential at \(i\):

\begin{equation}\phantomsection\label{eq-accesssibility-01}{
V_{i} = k \sum_j W_j^{(2)} f(c_{ij})
}\end{equation}

Similar to Eq~\ref{eq-phys-gravity-model}, \(W_j^{(2)}\) above is the
mass at the destination and the sub-indices are for \(i = 1,s, n\)
origins and \(j = 1,s, m\) destinations.

The market potential variant can also be generally defined, which is
effectively the transpose of \(i\) and \(j\) in Eq~\ref{eq-access-01}
and Eq~\ref{eq-accesssibility-01} as follows:
\begin{equation}\phantomsection\label{eq-market-01}{
M_{ji} = k W_i^{(1)} f(c_{ij})
}\end{equation}

\begin{equation}\phantomsection\label{eq-market-potential-01}{
M_{j} = k \sum_j W_i^{(1)} f(c_{ji})
}\end{equation}

Similar to Eq~\ref{eq-phys-gravity-model}, \(W_j^{(2)}\) above is the
mass at the destination and the sub-indices are for \(i = 1,s, n\)
origins and \(j = 1,s, m\) destinations.

To detail the anatomy of \(V_{ij}\) and \(M_{ji}\) along with the
partial sums of \(V_i\) and \(M_j\), Figure
Fig~\ref{fig-analytical-device-conc-accessibility} illustrates the
accessibility analytical framework we propose using a simple 3 zone
system. Each measure's most disaggregate value is \(X_{ij}\), potential
for spatial interaction from \(i\) to \(j\). The \(X\) is a stand in for
the \(ij\) values of all the cases and their variants that will be
described (i.e.,
\(V_{ij}^0, M_{ji}^0, V_{ij}^T, M_{ji}^T, V_{ij}^S, M_{ji}^S\) and
\(V_{ij}^D, M_{ji}^D\)). The single marginals represent the origin-side
and destination-side weights of the zones. The total marginal represent
the sum of a single marginal.

\begin{figure}[H]

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{../figures/access-analytical-device.png}

}

\caption{\label{fig-analytical-device-conc-accessibility}The family of
accessibility measures analytical framework: labelling and associating
ij flows, zonal weights, the single marginals, and the total marginal.}

\end{figure}%

As an additional overview, we define four distinct members of the
constrained accessibility measure family, all delineated based on their
constant \(k\), which takes the form of either balancing factors
\(K^T\), \(B_j\), \(A_i\) depending on the indicator:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{The unconstrained case} with variants \(V_i^0\) and \(M_i^0\).
  This case is equivalent to Hansen's \citep{hansen1959} formulation
  (first variant) and Reilly's \citep{reilly1929methods} market
  potential formulation (second variant), whereby a balancing factor is
  neglected, so units of zonal values are in units of
  `opportunities-by-travel-impedance-value' and
  `population-by-travel-impedance-value'. In this case, no constraints
  are introduced to ensure that values of marginals are preserved.
\item
  \textbf{The total constrained case} with variant \(V_i^T\) and variant
  \(M_j^T\). The first variant \(V_i^T\)--the \emph{total constrained
  accessible opportunities measure}--resembles Hansen's
  \citep{hansen1959} formulation but with an additional regional
  balancing factor \(K^T\) term, defined as the ratio of the total
  number of opportunities in the region to the total sum of
  unconstrained accessibility values in the region. In this way, \(K^T\)
  ensures that each zonal accessibility value is a proportion of the
  total opportunities in the region (i.e., the total marginal in
  Fig~\ref{fig-analytical-device-conc-accessibility}), requires no
  information about the population seeking opportunities, and is in
  units of `opportunities' accessible. The second variant \(M_j^T\) is
  the transpose of \(i\) to \(j\) of the first variant, effectively a
  constrained version of market potential and is referred to as
  \emph{total constrained accessible population measure}. In this
  variant, each zonal accessibility value is a proportion of the total
  population in the region (maintained by \(\hat K^T\)), requires no
  information about the opportunities that are sought, and each zonal
  value is in units of `population' accessible.
\end{enumerate}

\begin{itemize}
\tightlist
\item
  When conceptualising \emph{potential} for spatial interaction--whether
  with opportunities or population--the total constrained case reflects
  the lowest level of restriction and hence maximum potential. The
  balancing factors \(K^T\) and \(\hat K^T\) only ensure that
  \(V_{ij}^T\) and \(M_{ji}^T\) values end up matching the \emph{total
  sum} of one of the single marginals, not the individual marginals
  themselves. For example, \(K^T\) does not guarantee that all the
  opportunities accessibility at \(V_{i=1}^T\) reflect a proportional
  allocation of the destination mass (i.e., \emph{number} of
  opportunities) at \(j=1, 2, 3\). In some cases, an allocation of
  opportunities from a destination \(j\) to \(i=1\) could \emph{exceed}
  the number of opportunities at that \(j\) (i.e., meaning that
  destination \(j\) is very attractive and reachable for an \(i\),
  relative other flows in the region). However, what \(K^T\) \emph{does}
  ensure is that each \(V_{ij}^T\) does not exceed the overall number of
  opportunities in the region--the total marginal. In other words,
  \(V_{i=1}^T\) expresses the number of opportunities that origin
  \(i=1\) could potentially interact with, as drawn from the entire
  regional opportunity mass, rather than constrained by individual
  destination totals. In this way, the \emph{total constrained} case
  reflects the maximum amount of \emph{potential} while still
  maintaining interpretable units.
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  \textbf{The singly constrained case}, with two variants \(V_i^S\)
  (opportunity-constrained) and \(M_j^S\) (population-constrained). The
  first variant is mathematically equivalent to the spatial availability
  measure \citep{soukhovIntroducingSpatialAvailability2023}, and this
  variant's per capita form is equivalent to the popular 2SFCA measure
  \citep{luo2003, shen1998}. Both variants can also be defined using
  either Hansen's \citep{hansen1959} or the market potential formulation
  but with balancing factor \(B_j\) for \(V_i^S\) or \(A_i\) for
  \(M_j^S\). These balancing factors ensure that the total marginal
  (green box in Fig~\ref{fig-analytical-device-conc-accessibility}) is
  maintained \emph{as well as} the values at one of the single marginals
  (destination-mass marginal for opportunity-constrained and origin-mass
  marginal for population constrained in
  Fig~\ref{fig-analytical-device-conc-accessibility}). In this way, the
  singly constrained case reflects a medium amount of potential,
  restricted by either values fitting the single destination-mass
  marginal or the single origin-mass marginal.
\end{enumerate}

\begin{itemize}
\tightlist
\item
  The first variant \(V_i^S\) includes a set of destination-side
  balancing factors \(B_j\) which ensure opportunities (or destination
  masses) are allocated to each origin \(i\) based on the population
  (i.e, origin mass) of the origin \(i\) and travel impedance from that
  \(j\) to all \(i\)s. Hence, \(B_j\) ensures that each \(V_i^S\) is the
  product sum of a \emph{proportional share} of opportunities allocated
  from each destination \(j\) implicitly also being a share of the total
  opportunities in the region. Similar to the total constrained case,
  each \(V_i^S\) and \(V_ij^S\) accessibility value is expressed in
  units of `opportunities' accessible. But unlike the total constrained
  case, the singly constrained case explicitly incorporates population
  by allocating a balanced share of a destination's total opportunities
  to populations at reachable origins.
\item
  The second variant \(M_j^S\) includes a set of opportunity-side
  balancing factors \(A_i\) which ensure population (or origin masses)
  are allocated to each destination \(j\) based on the destination mass
  of the destination \(j\) and all possible travel impedance from that
  \(i\) to all \(j\)s. \(A_i\) ensures similar, but transposed, so
  constraints are satisfied. Specifically, each \(M_j^S\) represents
  both a share of the total regional population and the sum of balanced
  proportions of population from all origins, allocated to each
  destination based on the opportunities sought and travel impedance.
  Each zonal value is expressed in units of `population' accessible.
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  \textbf{The doubly constrained case}. It is constrained simultaneously
  by population (origin masses) and opportunities (destination masses),
  so each \(V_{ij}^D\) (equivalent to \(M_{ji}^D\)) is expressed in
  units of `population-opportunity capacity' that is accessible between
  \(i\) to \(j\). This case requires the number of opportunities and
  population to match, e.g., the analyst must know the matching spatial
  interaction capacity of the population (demand) and opportunities
  (supply). By using this one-to-one matching data and the double
  constraints (i.e., \(A_i\) and \(B_j\) at once, ensuring both the
  double constraint maintains both marginals in
  Fig~\ref{fig-analytical-device-conc-accessibility}), this case
  restricts the \emph{potential} to spatially interact completely. In
  other words, the doubly constrained accessibility values reflect the
  number of predicted interactions between the opportunities and
  population, effectively, this case is equivalent formulaically as
  Wilson's doubly constrained spatial interaction model (i.e.,
  \emph{attraction-production constrained}). It can be understood as
  predicting a value of `access', and not accessibility.
\end{enumerate}

And as a summary: each member of the family of accessibility measure is
named, explained in plain language, alongside their balancing factor(s),
proportional allocation factor(s), and mathematical equation and value
interpretations in Table 1.

\blandscape

\input{summary-family-access-measures-table.tex}

Table 1: Summary of constrained accessibility measure types and
interpretations

\elandscape

\subsection{Setup of the simple numeric
example}\label{setup-of-the-simple-numeric-example}

Consider a simple region as shown in
Fig~\ref{fig-analytical-device-conc-accessibility}, with zone IDs 1, 2
and 3. Each zone is both an origin \(i\) and a destination \(j\). The
following three pieces of information are defined: zonal population and
opportunities, zonal cost matrix, and travel impedance functions for
three types of travel behaviour. In this example, the population are
people, the opportunities are physicians, and the region can be
described by three possible travel behaviours.

Firstly, Table~\ref{tbl-small-system-land-use} summarises the population
(in units of 10,000s of people) and the opportunities (the number of
physicians) per zone. For reference, considering
Table~\ref{tbl-small-system-land-use}'s values, the
Provider-to-Population-Ratio (PPR) in this system is 24.5. The number of
physicians per 10,000 in Canada in 2022 was 24.97
\citep{whoMedicalDoctors102025}. Secondly, to pair the zonal population
and opportunity information, the assumed cost of movement (in minutes of
travel time) between origins and destinations is as shown in
Table~\ref{tbl-small-system-cost}.

\begin{table}

\caption{\label{tbl-small-system-land-use}Simple system with three zones
(ID 1, 2 and 3). Population is in 10,000 persons and opportunities in
number of physicians.}

\centering{

\fontsize{7.5pt}{9.0pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}rcc}
\toprule
ID (i or j) & Population\textsuperscript{\textit{1}} & Opportunities\textsuperscript{\textit{2}} \\ 
\midrule\addlinespace[2.5pt]
1 & 4 & 160 \\ 
2 & 10 & 150 \\ 
3 & 6 & 180 \\ 
\bottomrule
\end{tabular*}
\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}Population is \emph{Wi\textsuperscript{(1)}} when used as a proxy for the mass at the origin, and \emph{Oi} when used as a constraint.\\
\textsuperscript{\textit{2}}Opportunities is \emph{Wj\textsuperscript{(2)}} when used as a proxy for the mass at the destination, and \emph{Dj} when used as a constraint.\\
\end{minipage}

}

\end{table}%

\begin{table}

\caption{\label{tbl-small-system-cost}Cost matrix for system with three
zones (travel time in minutes).}

\centering{

\fontsize{7.5pt}{9.0pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}cccc}
\toprule
 & \multicolumn{3}{c}{Destination ID} \\ 
\cmidrule(lr){2-4}
Origin ID & 1 & 2 & 3 \\ 
\midrule\addlinespace[2.5pt]
1 & 10 & 30 & 15 \\ 
2 & 30 & 10 & 25 \\ 
3 & 15 & 25 & 10 \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

From both Table~\ref{tbl-small-system-land-use} and
Table~\ref{tbl-small-system-cost}, Zone 3 and 1 can be interpreted as of
an urban core with major healthcare institutions and a healthcare
cluster on the edge of the city respectively, each with a moderate
residing population (with zone 3 having both a higher population and
physician count). Zone 2, can be interpreted as a more distant bedroom
community, with a relatively high population and fewer physicians. In
sum, Zones 1 and 3 are more proximate to each other than to Zone 2, and
together match Zone 2's population while offering more than twice the
physician availability.

And lastly, in Eq~\ref{eq-travel-behaviour-scenarios} we distinguish
accessibility measure values for the following three impedance functions
that represent the potential for spatial interaction travel behaviour of
the population to opportunities. Accessibility will be calculated three
times for each case, one assuming the most decay (\(f_1(c_{ij})\)),
another assuming medium decay (\(f_2(c_{ij})\)), and a third assuming
the least decaying travel behaviour (\(f_3(c_{ij})\)) for the entire
region. A helpful analogy may be tying travel behaviour to the used
mode's mobility potential, i.e., the most decaying travel behaviour
(\(f_1(c_{ij})\)) would assume all travel in the region being done by
foot, while calculating accessibility assuming the least decay
(\(f_3(c_{ij})\)) would assume unfettered automobility. Or
alternatively, these functions could represent travel behaviour on
snowstorm-affected day (\(f_1(c_{ij})\)) for the entire region versus a
clear, ideal travel day (\(f_3(c_{ij})\)). As an example of a discussion
on how travel behaviour has been considered in accessibility measures
cost of travel see Páez et al. \citep{paez2012measuring}.

\begin{equation}\phantomsection\label{eq-travel-behaviour-scenarios}{
\begin{array}{l}
f_1(c_{ij}) = \frac{1}{c_{ij}^3}\\
f_2(c_{ij}) = \frac{1}{c_{ij}^2}\\
f_3(c_{ij}) = \frac{1}{c_{ij}^{0.1}}
\end{array}
}\end{equation}

Any set of concepts representing population, opportunities, and their
associated travel behaviour, whether representing the entire region
uniformly (as we will demonstrate) or representing specific subgroups,
can be substituted into our simple example, depending on the research
question. The purpose of the following simple example is to demonstrate
the calculation of each member of the accessibility measure family,
interpret the values, and compare them both within and across travel
behaviour groups and members of the family of accessibility measures.

\subsection{Unconstrained
accessibility}\label{unconstrained-accessibility}

Setting the balancing factor \(k\) to 1 or omitting it completely in
Eq~\ref{eq-access-01} results in the unconstrained accessibility case:

\begin{equation}\phantomsection\label{eq-unconstrained-access}{
V^0_{ij} = 1 *W_j^{(2)} f(c_{ij})
}\end{equation}

In this case, the partial sum of spatial interaction is simply identical
to Hansen's accessibility \(S_i\) \citep{hansen1959}, the current
standard practice in accessibility measurement:

\begin{equation}\phantomsection\label{eq-unconstrained-accessibility}{
V^0_i = \sum_j V^0_{ij} = \sum_j W^{(2)}_jf(c_{ij}) = S_i
}\end{equation}

The sum of the unconstrained accessibility values for each origin
\(V^0_{i}\) generally does not equal the total number of opportunities
\(O\) (e.g., \(\sum_i V^0_{i} \not= O\)), since arbitrarily setting
\(k\) to 1 (or neglecting \(k\) all together) strips the values of any
meaningful unit-based interpretation, the units are in `summed
opportunities by some travel impedance value'. Moreover, comparisons of
these \(V^0_{i}\) values across different contexts such as different
impedance functions \(f(c_{ij})\) or varying number of zones exacerbates
this issue as the units between \(V^0_{i}\) values change i.e.,
comparisons between a value of `summed opportunities by a travel
impedance value' to a value of `summed opportunities by another travel
impedance value' are not directly interpretable. From this perspective,
the raw unconstrained accessibility scores are not intuitively
comparable across different contexts and decay functions. They should
more appropriately be used as an ordinal variable to make comparisons of
size (i.e., greater than, less than, equal to), not to calculate ratios
or intervals (i.e., the magnitude of differences).

Returning to our numeric example, the calculated unconstrained
accessibility \(V^0_{i}\) for each origin, a sum of all the travel
impedance weighted opportunities at each destination
(\(\sum_i V^0_{i}\)), in displayed in
Table~\ref{tbl-simple-example-unconstrained-accessibility}.

\begin{table}

\caption{\label{tbl-simple-example-unconstrained-accessibility}Simple
system: unconstrained accessibility.}

\centering{

\fontsize{6.0pt}{7.2pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}l|ccc}
\toprule
 & \multicolumn{3}{c}{V\textsubscript{i}\textsuperscript{0}} \\ 
\cmidrule(lr){2-4}
 & f\textsubscript{1} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{3} & f\textsubscript{2} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{2} & f\textsubscript{3} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{0.1} \\ 
\cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(lr){4-4}
Origin & units: \emph{physicians-minute\^{}-3} & units: \emph{physicians-minute\^{}-2} & units: \emph{physicians-minute\^{}-0.1} \\ 
\midrule\addlinespace[2.5pt]
1 & 0.219 & 2.567 & 371.143 \\ 
2 & 0.167 & 1.966 & 363.479 \\ 
3 & 0.237 & 2.751 & 373.738 \\ 
\midrule 
\midrule 
Sum & 0.6233422 & 7.283556 & 1108.361 \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

As the different impedance functions represent different travel
behaviours, comparing the raw unconstrained accessibility values across
groups is meaningless beyond notions of higher or lower. For instance,
at zone 1 the difference between the least decay (\(f_3(c_{ij})\)) and
most decay (\(f_1(c_{ij})\)) groups is 370.92, but in what units? These
two values are a product of different impedance functions, making the
comparison uninterpretable in absolute terms. Likewise, we could compare
values within the same travel behaviour scenario across different zones,
as they are in the same units, however the issue of unit
interpretability will also be apparent. Considering the most decaying
scenario \(f_1(c_{ij})\) and zone 1 (the zone with a healthcare cluster
at the edge of the city): zone 1 captures 0.0181185 fewer
physicians-minute\(^{-1}\) than zone 3 (urban core). Again, the
fundamental uninterpretability of what is a
\emph{physicians-minute\(^{-1}\)} or
\emph{opportunity-weighted-travel-impedance} unit remains.

While one could attempt to adjust the units post-calculation (e.g.,
scaling, population normalization) or select impedance functions to
facilitate comparison across scenarios (potentially at the expense of
accurately reflecting travel behavior), such adjustments may introduce
bias. Hence, the raw unconstrained accessibility values themselves are
challenging to compare due to their units. To enable more meaningful
comparison, the following sections will detail the introduction of
constraining constants to ensure consistent units across scenarios and
demonstrate results on the same numeric example.

\subsection{Total constrained
accessibility}\label{total-constrained-accessibility}

In the total constrained accessibility case, a total balancing factor
proportionally adjusts unconstrained zonal accessibility values
\(V^0_i\) based on the regional sum of \(V^0_i\) and the total
population or opportunities in the region. Alternatively, we reformulate
this case using a proportional allocation constant, which allocates
opportunities (or population) proportionally based on the the travel
impedance and the total population or opportunities in the region. In
both formulations, all zonal values become a proportion of a known
system total, be it the regional opportunities or regional population
depending on the variant.

We define two variants for this case: (a) \(V_i^T\) where accessibility
is constrained by the total number of opportunities (total constrained
accessible opportunity) and which is interpreted as Hansen's
accessibility with a constraining constant, and (b) \(M_j^T\), where
\(i\) and \(j\) of the first variant is transposed, yielding a measure
constrained by the total number of population and to be interpreted as
constrained `market potential'.

\subsubsection{Total constrained accessible opportunities: Hansen's
accessibility with a total
constraint}\label{total-constrained-accessible-opportunities-hansens-accessibility-with-a-total-constraint}

Unlike in Eq~\ref{eq-access-01}, the proportionality constant \(k\) is
retained. For the total constrained case, it is represented as \(K^T\):

\begin{equation}\phantomsection\label{eq-total-constrained-access}{
V^T_{ij} = K^T  W_j^{(2)}  f(c_{ij})
}\end{equation}

In this way, the total constrained accessibility measure now becomes
Hansen's accessibility with a balancing factor \(K^T\):

\begin{equation}\phantomsection\label{eq-total-constrained-accessibility}{
V^T_i = \sum_j V^T_{ij} = K^T \sum_j W^{(2)}_jf(c_{ij}) = K^T  V^0_i
}\end{equation}

Imagine that the only system known is the total number of opportunities
\(D\) in the region. Accordingly, the constant we impose in this case
ensures the regional sum of total constrained accessibility is equal to
the total number of opportunities as follows:
\begin{equation}\phantomsection\label{eq-total-constraint-accessibility}{
\sum_i V^T_{i} = \sum_i\sum_j V^T_{ij} = D
}\end{equation}

This constraint is analogous to the total constraint of
Eq~\ref{eq-constraint0-gravitymodel}, congruent with Wilson's framework.
Given the total number of opportunities in the region, we can then
substitute Eq~\ref{eq-total-constrained-accessibility} in
Eq~\ref{eq-total-constraint-accessibility} to solve for \(K^T\):

\begin{equation}\phantomsection\label{eq-total-opportunity-balancing-factor}{
K^T =\frac{D}{\sum_i\sum_j V^0_{ij}} = \frac{D}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
}\end{equation}

Which is also congruent with Wilson's framework as it comparable to the
total flow spatial interaction model (e.g., Equation 2.11 in
\citep{cliff_evaluating_1974}). Hence, rearranging the equation to have
opportunities and the proportional constant distinctly represented, our
total constrained accessibility model is:

\[
V^T_i = K^T\sum_j W^{(2)}_jf(c_{ij}) = \sum_j D\frac{W^{(2)}_j f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
\]

Further, we can see that, since \(D\) and \(W^{(2)}_j\) are both in
units of opportunities, the proportional allocation factor for the total
constrained opportunity case \(\kappa_{ij}^T\) is dimensionless:

\[
\kappa_{ij}^T = \frac{W^{(2)}_j f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
\]

\noindent and therefore \(V^T_i\) is now in the units of \(D\), that is,
the mass at the destination (\(V^T_i = \sum_j \kappa_{ij}^T D\)). The
role of \(\kappa_{ij}^T\) in this reformulation of accessibility is to
transform between units and to adjust the number of opportunities
accessible from \(i\) so they represent a proportion of the total number
of opportunities in the region \(D\). \(\kappa_{ij}^T\) then assigns
opportunities in proportion to the impedance between \(i\) and \(j\).
This is why we refer to \(\kappa_{ij}^T\) as a proportional allocation
factor. On the other hand, the proportionality constant \(K^T\) balances
the units of \(V^0_i\), the Hansen-type accessibility values, and is an
alternative expression of the total constrained accessibility measure.

Referring back to our simple numeric example, balancing factor \(K^T\)
for the most decay travel behaviour scenario
\(f_1(c_{ij}) = 1/c_{ij}^3\) would then be:

\[
\begin{array}{l}
K^T = \frac{D}{\sum_{i}\sum_{j} W_j^{(2)} f(c_{ij})}\\
K^T = \frac{D}{\frac{W_1^{(2)}}{c_{11}^3}+\frac{W_1^{(2)}}{c_{21}^3} + \frac{W_1^{(2)}}{c_{31}^3} + s + \frac{W_3^{(2)}}{c_{31}^3} + \frac{W_3^{(2)}}{c_{32}^3} + \frac{W_3^{(2)}}{c_{33}^3}
}\\
K^T = \frac{490}{0.6233422} \\
K^T = 786.085
\end{array}
\]

Using the calculated proportionality constants for all zones and
multiplying them by the unconstrained accessibility value \(V^0_i\), the
total opportunity constrained accessibility values for all zones and
different travel behaviour scenarios is presented in
Table~\ref{tbl-simple-example-total-opportunity-accessibility}.
\(\kappa_{ij}^T\) for each zone is not shown, but can be understood to
be the unitless proportion of opportunities (of the total opportunities)
allocated to each zone based on travel impedance.

\begin{table}

\caption{\label{tbl-simple-example-total-opportunity-accessibility}Simple
system: total constrained accessible opportunities.}

\centering{

\fontsize{6.0pt}{7.2pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}l|ccc}
\toprule
 & \multicolumn{3}{c}{V\textsubscript{i}\textsuperscript{T}} \\ 
\cmidrule(lr){2-4}
 & f\textsubscript{1} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{3} & f\textsubscript{2} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{2} & f\textsubscript{3} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{0.1} \\ 
\cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(lr){4-4}
Origin & units: \emph{physicians} & units: \emph{physicians} & units: \emph{physicians} \\ 
\midrule\addlinespace[2.5pt]
1 & 172.065 & 172.672 & 164.080 \\ 
2 & 131.627 & 132.247 & 160.692 \\ 
3 & 186.308 & 185.081 & 165.228 \\ 
\midrule 
\midrule 
Sum & 490 & 490 & 490 \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

In contrast to unconstrained accessibility \(V^0_i\), imposing a
constraint -in this case a total opportunity constraint for this
variant- allows for the comparison of differences and ratios between
regions and across different travel behaviour scenarios as well. Each
value is effectively in units of physicians, with the impedance units
already accounted for by \(\kappa_{ij}^T\).

Considering the highest decay scenario (\(f_1(c_{ij})\)), zone 1 (a
healthcare cluster at the edge of the city) captures an intermediate
amount of physicians (172.0652825) like in the unconstrained
accessibility case. However, unlike in the unconstrained case, we can
say that this value is out of the 490 physicians in the region, which
allows us also to deduce that zone 1 captures 1.3072213 and 0.9235529
times more than zone 2 and 3. Values for the lesser decay
(\(f_2(c_{ij})\)) and lowest decay (\(f_3(c_{ij})\)) scenarios are
calculated separately, with decay scenario values also summing to equal
490 physicians accessible in the region.

One can also directly compare values at a specific zone due to the
consistent units. For instance, zone 1 remains intermediate in capturing
accessible physicians relative to zones 2 and 3 across scenarios,
similar to the unconstrained case. However, the difference between
travel behaviour scenarios differ in direction. Specifically, Zone 1
captures 0.6067946 more and 7.9850478 fewer opportunities than the
lesser decay scenarios \(f_2(c_{ij})\) and \(f_3(c_{ij})\) respectively.
Why? \(\kappa_{ij}^T\) ensures proportional allocation for each travel
behaviour scenario. Meaning, while the unconstrained accessibility
increases, \(\kappa_{ij}^T\) adjusts the values to remain proportional
to the total number of opportunities (490 physicians accessible in the
region). As the decay behaviour decreases, more opportunities are
accessible for all zones. In the medium decay scenario \(f_2(c_{ij})\),
zone 1 sees a slight increase in values (relative to the highest decay
scenario) as the zone can accessible more opportunities relative to
increases seen in other zones. However, in the lowest decay scenario,
zone 1 sees a decrease, as it is outpaced by increases in other zones -
namely zone 2 (recall: zone 2 has the lowest number of opportunities,
hence the increases in opportunity gains is much higher in a low decay
scenario).

Using the total opportunity constrained formulation of accessibility
offers a solution to the unit interpretability issue of Hansen's
\citep{hansen1959} accessibility measure. Intuitively, the use of the
constraint illustrates how the differences and ratios of values between
zones and decay groups can be compared. This is true for other
constrained cases of the family of accessibility measures.

\subsubsection{Total constrained accessible population: Reilly's
potential trade territories with a total
constraint}\label{total-constrained-accessible-population-reillys-potential-trade-territories-with-a-total-constraint}

Another variant of the total constrained accessibility measure is the
\emph{total constrained accessible population measure}, which represents
the transpose of \(i\) to \(j\) of the total constrained accessible
opportunities measure. This variant, expressed in
Eq~\ref{eq-total-constrained-market}, represents an expression of the
concept of market potential (i.e., potential users) as proposed in
Harris \citep{harris_market_1954} and Vickerman
\citep{vickermanAccessibilityAttractionPotential1974}, and which Reilly
earlier referred to as `potential trade territories'
\citep{reilly1929methods}. This unconstrained form of market potential
\(M_j^0\) (Eq~\ref{eq-unconstrained-market}), effectively the \(i\) to
\(j\) transpose of \(V^0_{ij}\), has been used in recent research to
express the potentially accessible population (i.e., users) as a result
of regional transportation infrastructure investment projects
\citep[e.g.,][]{gutierrezLocationEconomicPotential2001, holl2007twenty, condecco2018road}.
Put another way, market potential can also be thought of as a form of
\emph{passive accessibility}, indicating the number of people that can
reach each destination. However, like \(V_{ij}^0\), issues of unit
interpretability arise in \(M_j^0\)'s unconstrained form. To address
this, the constrained variant, the total constrained accessible
population measure \(M^T_{j}\), is introduced. To formulate this
variant, the total balancing factor \(K^T\) is applied to the mass of
the \emph{population} at \(i\) (\(W_i^{(1)}\)) instead of the
opportunities at \(j\).

\begin{equation}\phantomsection\label{eq-total-constrained-market}{
M^T_{ji} = \hat K^T  W_i^{(1)} f(c_{ij}) = \hat K^T  M_{ji}^0
}\end{equation}

\noindent with \(M_j^0\) being the \(i\) \(j\) transpose of
\(V^0_{ij}\):

\begin{equation}\phantomsection\label{eq-unconstrained-market}{
M_j^0 = W_i^{(1)} f(c_{ij})
}\end{equation}

The total constrained accessible population measure (market potential)
then becomes:

\[
M^T_j = \sum_i M^T_{ji} = \hat K^T \sum_i W^{(1)}_if(c_{ij})
\]

Where we impose the total system known as a constraint, i.e., that the
total market potential equals the total population \(O\) in the region:

\begin{equation}\phantomsection\label{eq-total-constraint-market}{
\sum_j M^T_{j} = \sum_i\sum_j M^T_{ji} = O
}\end{equation}

Substituting Eq~\ref{eq-total-constrained-market} in
Eq~\ref{eq-total-constraint-market}, and solving for \(\hat K^T\), we
obtain:

\begin{equation}\phantomsection\label{eq-total-population-balancing-factor}{
\hat K^T=\frac{O}{\sum_i\sum_i M^T_{ji}} = \frac{O}{\sum_i\sum_j W^{(1)}_i f(c_{ij})}
}\end{equation}

The constrained market potential then takes the following form: \[
M^T_j = \hat K^T \sum_i W^{(1)}_if(c_{ij}) = O \sum_i \frac{W_i^{(1)} f(c_{ij})}{\sum_i\sum_j W^{(1)}_i f(c_{ij})}
\]

Where the following \(\hat\kappa_{ji}^T\) proportional allocation factor
is dimensionless, reflecting the proportional allocation of population
in the region \(O\) (i.e., \(M^T_j = \sum_i \hat \kappa_{ji}^T O\)) \[
\hat \kappa_{ji}^T = \frac{W^{(2)}_i  f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
\]

Returning back to the numerical example, the balancing factor
\(\hat K^T\) is solved for each travel behaviour scenario, and the
market potential of each zone \(M^T_j\) is expressed as units of
population (e.g., the number of people accessible from each origin at
that destination) in
Table~\ref{tbl-simple-example-total-population-accessibility}.

\begin{table}

\caption{\label{tbl-simple-example-total-population-accessibility}Simple
system: total constrained accessible population.}

\centering{

\fontsize{6.0pt}{7.2pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}l|ccc}
\toprule
 & \multicolumn{3}{c}{M\textsubscript{i}\textsuperscript{S}} \\ 
\cmidrule(lr){2-4}
 & f\textsubscript{1} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{3} & f\textsubscript{2} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{2} & f\textsubscript{3} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{0.1} \\ 
\cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(lr){4-4}
Destination & units: \emph{population in 10,000s} & units: \emph{population in 10,000s} & units: \emph{population in 10,000s} \\ 
\midrule\addlinespace[2.5pt]
1 & 5.018 & 5.447 & 6.598 \\ 
2 & 8.596 & 7.986 & 6.717 \\ 
3 & 6.386 & 6.567 & 6.684 \\ 
\midrule 
\midrule 
Sum & 20 & 20 & 20 \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

Readers may note the difference in trends in accessible population
(Table~\ref{tbl-simple-example-total-population-accessibility}) and
accessible physicians (i.e., the preceding subsection,
Table~\ref{tbl-simple-example-total-opportunity-accessibility}). In
Table~\ref{tbl-simple-example-total-opportunity-accessibility}, zone 1,
2, 3 represent destinations and the accessibility values reflect the
number of accessible people from the vantage of physicians. Zone 1, in
its role as a destination, is no longer intermediately-ranked relative
to other zones; it now attracts the fewest number of people across all
three travel behaviour scenarios. However, similar to the total
constrained opportunity case, as travel decay reduces, the availability
of population begins to converge (though Zone 1 continues as the
lowest-ranked) for similar reasons. As decay reduces, the population's
travel impedance to all zones become more similar, making the relative
location of the zones less important and all people in the region more
equally accessible. Overall: like the total constrained accessible
opportunities case, this variant allows for the interpretation of both
ordinal and interval comparisons of the raw values themselves.

\subsection{Singly constrained
accessibility}\label{singly-constrained-accessibility}

Similar to the total constrained accessibility measure, the singly
constrained case includes a balancing factor that adjusts the
unconstrained zonal accessibility values \(V^0_i\) such that a
\emph{single} constraint is satisfied. Two variants are defined: (a) the
\emph{singly constrained accessible opportunities} case (an alternative
formula to spatial availability in Soukhov et al.
\citep{soukhovIntroducingSpatialAvailability2023}), and (b) the
\emph{singly constrained accessible population} case, its transpose, or
singly constrained market potential.

Unlike the total constraint (i.e.,
Eq~\ref{eq-total-constraint-accessibility}), the single constraint--as
will be defined in Eq~\ref{eq-opportunity-constraint} and
Eq~\ref{eq-population-constraint} for the first and second variants--
incorporates additional information. In the opportunities-accessible
variant, the associated balancing factor constrains the `potential' for
spatial interaction by ensuring that only a proportional amount of
opportunities at each destination are allocated to `demanding' origins.
This allocation is informed by demand, or the relative amount of
population, and associated travel impedance connecting the zones. In the
population-accessible variant (i.e., the \(i\) -\textgreater{} \(j\)
transpose of the first variant), population at each origin is
proportionally allocated to destinations based on the share of
opportunities and travel impedance.

In both variants, the singly constrained accessibility measure
introduces population-based (or opportunity-based) competition at the
zonal level, unlike the total constraint, which more simply allocates a
fixed regional total of opportunities (or population depending on the
variant). In sum, all singly constrained zonal accessibility values are
both a proportion of the known regional opportunity (or population)
total \emph{and} a sum of a \emph{balanced} proportion of opportunities
allocated from each destinations (or population allocated from each
origin). Each zonal value remains in units of opportunities accessible
(or population accessible).

\subsubsection{Singly constrained accessible opportunities: spatial
availability}\label{singly-constrained-accessible-opportunities-spatial-availability}

To demonstrate this variant formulaically, we begin with the opportunity
constraint (Eq~\ref{eq-opportunity-constraint}) as our known piece of
information. Namely, the sum of accessible opportunities from a
destination should equal the number of opportunities \(D_j\) at that
destination. As the number of opportunities at each \(j\) are known, it
is represented as \(D_j\) instead of \(W_j^(2)\) as in the total
constrained case. This constraint should hold for all destinations in
the region. This is comparable to the single attraction-constraint
(Eq~\ref{eq-constraint2-gravitymodel}) from Wilson's framework.

\begin{equation}\phantomsection\label{eq-opportunity-constraint}{
\sum_i V^S_{ij} =  D_j
}\end{equation}

The underlying spatial interaction model is now the
attraction-constrained model in
Eq~\ref{eq-attraction-constrained-gravitymodel}, and our accessibility
measure becomes:

\begin{equation}\phantomsection\label{eq-opportunity-constrained-accessibility}{
V^S_{i} = \sum_j B_j D_j W_i^{(1)} f(c_{ij})
}\end{equation}

\noindent where \(W_i^{(1)}\) is a measure of the mass at origin \(i\)
(i.e., the opportunity-seeking population). The corresponding balancing
factor, as per Wilson, is:

\begin{equation}\phantomsection\label{eq-opportunity-constrained-proportionality-constants}{
B_j = \frac{1}{\sum_i W_i^{(1)} f(c_{ij})}
}\end{equation}

Introducing the balancing factor in
Eq~\ref{eq-opportunity-constrained-accessibility}, we obtain:

\begin{equation}\phantomsection\label{eq-opportunity-constrained-accessibility-with-balancing-factor}{
V^S_{i} = \sum_j D_j \frac{W_i^{(1)} f(c_{ij})}{\sum_i W_i^{(1)} f(c_{ij})}
}\end{equation}

Further, we define the following proportional allocation factor:

\begin{equation}\phantomsection\label{eq-opportunity-constrained-proportional-allocation-factor}{
\kappa^S_{ij} = \frac{W_i^{(1)} f(c_{ij})}{\sum_i W_i^{(1)} f(c_{ij})}
}\end{equation}

After this, it is possible to rewrite
Eq~\ref{eq-opportunity-constrained-accessibility-with-balancing-factor}
as an origin summary expression of proportionally allocated known
opportunities (i.e., \(D_j\)).

\begin{equation}\phantomsection\label{eq-attraction-constrained-accessibility-with-proportional-allocation-factor}{
V^S_{i} = \sum_j \kappa^S_{ij} D_j
}\end{equation}

Soukhov et al. \citep{soukhovIntroducingSpatialAvailability2023} have
shown that the role of \(\kappa^S_{ij}\) is to allocate opportunities
\(D_j\) proportionally to the mass at each origin \(i\) and the
impedance between \(i\) and \(j\). As in the total constrained
opportunity case, \(\kappa^S_{ij}\) is dimensionless and \(V_i^S\) is in
the units of opportunities \(D_j\). The singly constrained accessibility
measure in
Eq~\ref{eq-attraction-constrained-accessibility-with-proportional-allocation-factor}
is called spatial availability by Soukhov et al.
\citep{soukhovIntroducingSpatialAvailability2023}, because it represents
the number of opportunities that can be reached \emph{and} are
available, in the sense that accessible opportunities have been
proportionally allocated based on relative demand, travel impedance and
the regional total number of opportunities, i.e., spatial competition
for them has been considered. These authors also show that the following
expression (accessibility per capita) is a constrained version of the
popular two-stage floating catchment area measure of Shen
\citep{shen1998} and Luo and Wang \citep{luo2003}:

\begin{equation}\phantomsection\label{eq-opportunity-constrained-accessibility-per-capita}{
v^S_{i} = \frac{V^S_{i}}{W^{(1)}_i}
}\end{equation}

Returning to the simple numeric example, the opportunity-constrained
case would yield the following \(B_{j}\) for \(f_1(c_{ij})\):

\[
\begin{array}{l}
B_{j} = \frac{1}{\sum_i W_i^{(1)} f(c_{ij})}\\
B_{1} =  \frac{1}{\frac{4}{10^3} + \frac{10}{30^3} + \frac{6}{15^3}} = 162.6506\\ 
B_{2} =  \frac{1}{\frac{4}{30^3} + \frac{10}{10^3} + \frac{6}{25^3}} = 94.9474\\
B_{3} =  \frac{1}{\frac{4}{10^3} + \frac{10}{25^3} + \frac{6}{10^3}} = 93.9850
\end{array}
\]

The balancing factors \(B_j\) for the \(f_2(c_{ij})\) decay group for
zones 1, 2 and 3 is 12.8571429, 8.7685113 and 10.6635071, and for
\(f_3(c_{ij})\) decay group is 0.0672461, 0.0660559 and 0.0663798. Using
these these balancing constants, we can calculate the singly constrained
opportunity accessibility
(Table~\ref{tbl-simple-example-attraction-constrained-accessibility}).

\begin{table}

\caption{\label{tbl-simple-example-attraction-constrained-accessibility}Simple
system: singly constrained accessible opportunities.}

\centering{

\fontsize{6.0pt}{7.2pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}l|cccc}
\toprule
 &  & \multicolumn{3}{c}{V\textsubscript{i}\textsuperscript{S}} \\ 
\cmidrule(lr){3-5}
 &  & f\textsubscript{1} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{3} & f\textsubscript{2} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{2} & f\textsubscript{3} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{0.1} \\ 
\cmidrule(lr){3-3} \cmidrule(lr){4-4} \cmidrule(lr){5-5}
Origin & Population (10k) & units: \emph{physicians} & units: \emph{physicians} & units: \emph{physicians} \\ 
\midrule\addlinespace[2.5pt]
1 & 4 & 133.469 & 122.255 & 98.848 \\ 
2 & 10 & 166.781 & 185.096 & 241.877 \\ 
3 & 6 & 189.750 & 182.650 & 149.275 \\ 
\midrule 
\midrule 
Sum & — & 490 & 490 & 490 \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

Imposing the single proportional allocation factor \(\kappa^S_{ij}\)
allows for the comparison of differences and ratios of the accessibility
values, like previously discussed in the total constrained accessible
opportunities case. The proportional allocation factor ensures that
resulting values are in units of \emph{physicians}, with the impedance
units already accounted for in the allocation process.

However, unlike the total constrained opportunity case,
\(\kappa^S_{ij}\) reflects zonal competition based on the mass of the
origin, i.e., population. Again, consider the highest decay scenario
\(f_1(c_{ij})\). Under this scenario, zone 1 no longer captures a medium
amount of physicians as in the total constrained opportunity case: it
now captures the fewest in the region i.e., 133.4687282 at zone 1,
166.7813387 at zone 2, and 189.7499331 at zone 3. Why may zone 1
(healthcare cluster at the edge of the city) capture 27\% of the
physicians regionally while this same zone captures 35\% in the total
opportunity constrained case?

This difference is due to the single-opportunity factor
\(\kappa^S_{ij}\)'s role in \(V_i^S\). The only inputs required in the
total-opportunity factor \(\kappa^T_{ij}\) is the total number of
opportunities in the region \(D\) as well as the associated
opportunities at each \(j\) and travel impedance. Opportunities are
allocated to \(i\)s, regardless of the mass weights of origin. Whereas
the single opportunity constraint \(\kappa^S_{ij}\) requires the
population at \(i\) as an input. In fact, \(\kappa^S_{ij}\) is
calculated as the proportion of impedance-weighted population at an
\(i\) to the sum of impedance-weighted population for the entire region.
Hence, since zone 1 has the lowest population in the region, is in close
proximity to a more populated zone (zone 3, the `urban core'), and is
not well connected (in terms of travel impedance) to other zones with
opportunities, \(V_{1}^S\) values, or the number of physicians
accessible, is lowest.

Readers may also notice the change in the proportion of opportunities
drawn from different zones depending on the travel behaviour scenarios
considered. For instance, consider Zone 2 which has the highest
population. It is more evident in the \(f_3(c_{ij})\) scenario than in
higher decay scenarios that this zone does not have an exceptionally
large population for the region - Zone 2 only represents 50\% of the
population in the three zone region. In this sense, other zones are not
\emph{that} disadvantaged, and in this scenario with unfettered travel
cost, Zones 1 and 3 also take opportunities from Zone 2 (i.e., Zone 1
and Zone 3 takes 6\% and 8\% more from Zone 2 between \(f_3(c_{ij})\)
and \(f_1(c_{ij})\) scenarios hence \(\kappa_{2,2}^S\) decreases by
14\%). Zones 1 and 3 are allocated opportunities at relates similar to
their relative population size.

Readers may also notice the change in the proportion of opportunities
drawn from different zones depending on the travel behaviour scenarios
considered. For instance, Zone 2 has the highest zonal population,
representing 50\% of the 200,000 regional population. However, due to
its high relative travel distance from the other zones, its population
is less competitive in capturing opportunities in the high-decay travel
scenario (\(f_1(c_{ij})\)): Under this scenario, zone 2 captures almost
exclusively opportunities from its own zone. However, in
\(f_3(c_{ij})\), the scenario with unfettered travel cost, Zone 2
captures by far the most number of physicians. But in this scenario,
Zones 1 and 3 also take opportunities from Zone 2 (i.e., Zone 1 and Zone
3 takes 6\% and 8\% more from Zone 2 between \(f_3(c_{ij})\) and
\(f_1(c_{ij})\) scenarios hence \(\kappa_{2,2}^S\) decreases by 14\%).
Zones 1 and 3 are allocated opportunities at rates similar to their
relative population size.

In this way, the consideration of constrained accessibility \emph{per
capita} may be clarifying. Often, accessibility values are reported as
raw scores without the consideration for population. But, as we
introduced constraints, these constrained accessibility values can be
normalized using anything that is relevant to the zone. In
Table~\ref{tbl-simple-example-attraction-constrained-accessibility-per-capita},
we present per capita accessibility for the numeric example, simply in
units of number of physicians accessible per population at each zone.
Notably, these per capita rates are equivalent to the 2SFCA values.

\begin{table}

\caption{\label{tbl-simple-example-attraction-constrained-accessibility-per-capita}Simple
system: singly constrained accessible opportunities per capita.}

\centering{

\fontsize{6.0pt}{7.2pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}l|cccc}
\toprule
 &  & \multicolumn{3}{c}{v\textsubscript{i}\textsuperscript{S}} \\ 
\cmidrule(lr){3-5}
 &  & f\textsubscript{1} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{3} & f\textsubscript{2} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{2} & f\textsubscript{3} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{0.1} \\ 
\cmidrule(lr){3-3} \cmidrule(lr){4-4} \cmidrule(lr){5-5}
Origin & Population (10k) & units: \emph{physicians per capita} & units: \emph{physicians per capita} & units: \emph{physicians per capita} \\ 
\midrule\addlinespace[2.5pt]
1 & 4 & 33.367 & 30.564 & 24.712 \\ 
2 & 10 & 16.678 & 18.510 & 24.188 \\ 
3 & 6 & 31.625 & 30.442 & 24.879 \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

This simple example was constructed so that the regional average equals
24.5 physicians per 10,000 people. As distance decay decreases and
becomes \emph{relatively} uniform (all zones can reach all zones), the
effect of population drives the proportional allocation of
opportunities. Consequently, per capita accessibility values begin to
stabilise to the regional per capita average (e.g., in the lowest
distance decay \(f_3(c_{ij})\), per capita values are all around 24
physicians accessible per capita).

This trend mirrors how the accessibility values in the total constrained
opportunity case stabilises to the \(V_i^T\) regional average (e.g., the
accessible opportunities allocated to each of the three zones approaches
a third of 490 physicians, or 163.33, under the unfettered mobility
scenario \(f_3(c_{ij})\)). These patterns make intuitive sense: the
balancing factors act as regional and/or zonal averaging mechanisms. In
distance decay travel behaviour scenarios that are more relatively
uniform (i.e., low for all zones like in \(f_3(c_{ij})\)), what remains
is the relative effect of the other variables in the balancing factor.
In the total constrained case, this is the proportion of opportunities
relative to the regional opportunities, and in the case of the single
opportunity constrained case, this is the population at a zone relative
to the regional population.

\subsubsection{Singly constrained accessible population: market
availability}\label{singly-constrained-accessible-population-market-availability}

Similar to Eq~\ref{eq-total-population-balancing-factor} in transposing
the origins and destinations, we can define a \emph{singly constrained}
measure of market potential that preserves the known population (i.e.,
the mass weight at the origin \(W_i^{(1)}\) is now represented by
\(O_i\)). In it's per-capita expression, i.e., equivalent to 2SFCA, this
constrained concept of market potential been used to express ``facility
crowdedness'' as in Wang \citep{wang_inverted_2018}.

The underlying spatial interaction model is now the
production-constrained model in
Eq~\ref{eq-production-constrained-gravitymodel}, and our market
potential measure \(M^S_j\) becomes:

\begin{equation}\phantomsection\label{eq-population-constrained-accessibility}{
M^S_j = \sum_i A_i O_i W_j^{(2)} f(c_{ij})
}\end{equation}

In this case, the measure is singly constrained by the population
\emph{by origin} (i.e., \(O_i\)), like
Eq~\ref{eq-constraint2-gravitymodel} from Wilson's framework:

\begin{equation}\phantomsection\label{eq-population-constraint}{
\sum_j M^S_{ji} =  O_i 
}\end{equation}

And the corresponding balancing factor, as per Wilson, is:

\begin{equation}\phantomsection\label{eq-population-constrained-proportionality-constants}{
A_i = \frac{1}{\sum_j W_j^{(2)} f(c_{ij})}
}\end{equation}

Following the same logic as in the preceding section on total
constrained market potential, one arrives at the following expression:

\begin{equation}\phantomsection\label{eq-production-constrained-accessibility-with-proportional-allocation-factor}{
M^S_{j} = \sum_i \hat \kappa^S_{ji} O_i
}\end{equation}

\noindent with:

\begin{equation}\phantomsection\label{eq-attraction-constrained-proportional-allocation-factor}{
\hat \kappa^S_{ji} = \frac{W_j^{(2)} f(c_{ij})}{\sum_i W_j^{(2)} f(c_{ij})}
}\end{equation}

As well, the single (population) constraint in
Eq~\ref{eq-population-constraint} ensures that the the total constraint
(e.g., \(\sum_j M^S_{j} = \sum_i\sum_j  M^S_{ji} = O\)) is maintained.

With these constraints, \(\frac{M_j^S}{O}\) can be interpreted as the
proportion of the total population serviced by location \(j\).

For the sake of brevity, we'll move forward onto the doubly constrained
case.

\subsection{Doubly constrained
accessibility}\label{doubly-constrained-accessibility}

This accessibility case requires zonal populations and opportunities to
match one-to-one, like the doubly constrained spatial interaction model.
In this way, doubly constrained accessibility can be thought as
``access'' or simply spatial interaction (no potential).

To contextualize this point, the total and singly constrained
accessibility measures discussed thus far have used either \(O_i\),
\(D_j\), or the regional sums of either, but never both simultaneously.
For example, when opportunities \(D_j\) are used to constrain
Eq~\ref{eq-opportunity-constrained-accessibility} in the \emph{singly
constrained accessible opportunities measure}, the specific mass of the
population at origin \(i\) demanding only those opportunities at that
\(j\) is unknown. Instead, only the population at \(i\) demanding
opportunities in the region is known, and this is more generally
represented as \(W_i^{(1)}\) (as in
Eq~\ref{eq-opportunity-constrained-proportionality-constants}).
Similarly, when the population \(O_i\) is used as a constraint in
Eq~\ref{eq-population-constrained-accessibility} in the \emph{singly
constrained accessible population measure}, the mass at the destination
is given by \(W_j^{(2)}\) (also in
Eq~\ref{eq-population-constrained-proportionality-constants}) since only
the mass of opportunities at each \(j\) is known and information on what
opportunities are allocated to what zone is unknown.

By contrast, the double-constrained accessibility case requires that
both populations and opportunities match. Meaning, opportunities at each
destination can be accessed by all populations, while at the same time,
the population at each origin can be accessed by all opportunities.
However, this requirement is often unintuitive in traditional
accessibility analysis. Namely, the distinction between population
(origin masses) and opportunities (destination masses) typically
represent different entities without a shared unit of measurement. On
the population side, we usually count people; on the opportunity side,
we may be referring to physicians, clinics, grocery stores, schools,
parks, or libraries. In a few cases, a one-to-one correspondence or
their own capacities at which they interact and are interacted with may
exist e.g., one person with one job. Another similar opportunity type
example is healthcare, e.g., one person and one unit of capacity, e.g.~a
vaccine shot.

A doubly constrained approach to accessibility calculation needs a
one-to-one relationships between population and opportunities to be
present. Mathematically, this model requires the simultaneous imposition
of both the population- and opportunity- constraints in the preceding
singly constrained variants (Eq~\ref{eq-opportunity-constraint} and
Eq~\ref{eq-population-constraint}), namely the sum of population in all
origins should match the sum of opportunities in all destinations
(Eq~\ref{eq-opportunity-population-equality}):

\begin{equation}\phantomsection\label{eq-opportunity-population-equality}{
\sum_i O_i = \sum_j D_j
}\end{equation}

As before, the simultaneous imposition of both constraints ensures the
total system constraint is maintained i.e., the sum of all doubly
constrained accessibility values
\(\sum_i V^D_{i} = \sum_i\sum_j  V^D_{ij} = D\) remains equal to the
total number of opportunities in the region \(O\) as shown in
Eq~\ref{eq-total-constraint-accessibility}.

As the doubly constrained accessibility measure \(V_{ij}^D\) takes the
form of the production-attraction (doubly constrained) spatial
interaction model, as shown in
Eq~\ref{eq-doubly-constrained-gravitymodel}, \(V_{i}^D\) is as follows:

\begin{equation}\phantomsection\label{eq-doubly-constrained-accessibility}{
V_{ij}^D = A_i B_j O_i D_j f(c_{ij})
}\end{equation}

\noindent where the corresponding balancing factors \(A_i\) and \(B_j\),
as per Wilson, are:

\[
\begin{array}{l}
A_i = \frac{1}{\sum_j B_j D_j f(c_{ij})}\\
B_j = \frac{1}{\sum_i A_i O_i f(c_{ij})}
\end{array}
\]

Calibration of the two sets of proportionality constants is accomplished
by means of iterative proportional fitting, whereby the values of
\(A_i\) are initialized as one for all i to obtain an initial estimate
of \(B_j\). The values of \(B_j\) are used to update the underlying
\(V_{ij}^D\) matrix, before calibrating \(A_i\). This process continues
to update \(A_i\) and \(B_j\) until a convergence criterion is met
\citep[see][p.~193-195]{ortuzar_2011_modelling}. The proportional
allocation factor \(\kappa_{ij}^D\) would then be:

\[
\kappa_{ij}^D = \sum_j \frac{1}{\sum_j B_j D_j f(c_{ij})} \frac{1}{\sum_i A_i O_i f(c_{ij})} O_i f(c_{ij})
\]

One could rewrite Eq~\ref{eq-doubly-constrained-accessibility} as an
origin summary expression of proportionally allocated opportunities:

\begin{equation}\phantomsection\label{eq-doubly-constrained-accessibility-w-paf}{
V^D_{i} = \sum_j \kappa^D_{ij} D_j
}\end{equation}

However, \(V^D_{i}\) is not wholly helpful, as it will simply equal the
origin-mass marginal. In this way, only \(V^D{ij}\) values are
interpretable. Furthermore, unlike in the total and singly constrained
cases, the doubly-constrained case does not have an interpretable per
capita version. For instance, representing \(V_{ij}^D\) per capita is
not meaningful, as the value \emph{already} matches population to
opportunities. Following this logic, the market potential form
\(M^D_{ij}\) is effectively equivalent to \(V_{ij}^D\), but can be read
with a different interpretation: i.e., the opportunities accessed from
\(j\) at an \(i\) vs.~the population accessed from \(i\) at a \(j\). The
inputs of `opportunities accessed' and `accessed population' can already
be interpreted as inherently being sensitive to both opportunities and
population.

To calculate doubly constrained accessibility, the interpretation of the
population data and the counts of the opportunity data in the numeric
example must be modified. Namely, a count of physician \emph{capacity}
per destination is needed instead of just the number of physicians, as
used to calculated total and singly constrained cases. We also must be
able to clearly state that the population is a count of people seeking
opportunities at the new capacities, i.e., the population must reflect
the \emph{capacity} of the population to interact with opportunities.

So, this adjusted simple example is summarised in
Table~\ref{tbl-small-system-land-use-doubly-constrained-case}: with the
population (in units of 10,000s of people seeking physicians) and the
opportunities (in units of 10,00s of physician-capacity) per zone. For
the population, we leave this unchanged numerically but theoretically
know that each person interacts with one physician capacity (i.e., our
opportunities with a \emph{capacity}). Hence, the number of
opportunities per destination is new: the example is modified such that
the physician-capacity at each zone is an approximately scaled version
of the number of destination-side physicians at each zone from the
unmodified example (Table~\ref{tbl-small-system-land-use}). To emphasis
the new definition of `provider' as physician capacity, the new system
PPR is simply 1, this is compared to the unmodified example which yields
system PPR of24.5. We keep the same zonal cost matrix, and travel
impedance functions for three types of travel behaviour as before
(Table~\ref{tbl-small-system-cost} and
Eq~\ref{eq-travel-behaviour-scenarios}).

\begin{table}

\caption{\label{tbl-small-system-land-use-doubly-constrained-case}Modified
simple system with three zones reflecting matched population and
opportunities. Population is in 10,000 persons and opportunities in
10,000 of physician-capacity.}

\centering{

\fontsize{7.5pt}{9.0pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}rcc}
\toprule
ID (i or j) & Population & Opportunities \\ 
\midrule\addlinespace[2.5pt]
1 & 4 & 7 \\ 
2 & 10 & 5 \\ 
3 & 6 & 8 \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

However, despite the modifications to the example, our objective remains
the same as in previous cases: to measure accessibility under different
travel behavior scenarios. Specifically, we aim to quantify the number
of \emph{potential} spatial interactions between the physician-seeking
population in each zone \(i\) to the physician capacity in a zone \(j\)
i.e., \(V_{ij}^D\). The highest decay travel behaviour scenario
(\(f_1(c_ij)\)) is presented in
Table~\ref{tbl-adjusted-small-system-land-use-doubly-constrained-case-f1cij-access-values}.

\begin{table}

\caption{\label{tbl-adjusted-small-system-land-use-doubly-constrained-case-f1cij-access-values}Doubly
constrained accessible opportunities assuming highest travel decay in
the modified simple system.}

\centering{

\fontsize{7.5pt}{9.0pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}l|rcccc}
\toprule
 &  & \multicolumn{3}{c}{Destination ID} &  \\ 
\cmidrule(lr){3-5}
 & Origin ID & 1 & 2 & 3 & sum \\ 
\midrule\addlinespace[2.5pt]
 & 1 & 3.235859 & 0.01032226 & 0.7556568 & 4 \\ 
 & 2 & 2.132602 & 4.95932483 & 2.9044391 & 10 \\ 
 & 3 & 1.631539 & 0.03035291 & 4.3399040 & 6 \\ 
\midrule 
\midrule 
Sum & — & 7 & 5 & 8 & — \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

As mentioned, accessibility values are typically interpreted as a
summary of the proportionally allocated opportunities at each \(i\).
Hence, in interpreting the doubly constrained accessibility from
Table~\ref{tbl-adjusted-small-system-land-use-doubly-constrained-case-f1cij-access-values},
\(V_i^D\) values (i.e., the sum of values at all three \(j\)
destinations for each origin \(i\)) would be 4.0018381, 9.9963656, and
6.0017963 physician-capacity accessible for Zones 1, 2 and 3
respectively. This approximately equal to the number of population at
each of these zones. Conversely, the market potential \(M_j^D\)
interpretation of these values would be 7, 5, and 8 people accessible
from Zones 1, 2 and 3 respectively, equal to the number of
\emph{opportunities} (physician-capacities accessible) at each of these
zones. Notice, the mass weight at the origin equals the mass weight of
the destination: this is precisely the function of the double
constraint. In other words, \(V_i^D\) is the number of accessed
opportunities and \(M_j^D\) is the number of population accessed. For
the other two travel behaviour, identical \(V_i^D\) and \(M_j^D\) values
are calculated, following the same logic. Hence, the usefulness of the
doubly constrained measure lies in the interpretation as \(V_{ij}^D\)
values.

For instance, differences in \(V_{ij}^D\) values between travel
behaviour scenarios are notable. These values can be directly compared
to discuss mass and distance decay impacts. Examining Zone 2 (the
bedroom community),
Table~\ref{tbl-adjusted-small-system-land-use-doubly-constrained-case-allfs-access-values-for-zone2}
demonstrates these \(i\) to \(j\) access values for this more relatively
remote, higher-populated and lower-opportunity rich zone. It can be
observed that the number of intrazonal opportunities proportionally
allocated decreases as the assumed distance decay decreases e.g., from
4.9593248 to 2.6672837 out of the \textasciitilde10 opportunities
allocated to Zone 2 (a population of 10). Following the intuition
discussed in the singly constrained opportunity case, as decay
decreases, the mass effect of the population (at origin) and opportunity
(at destination) is more evident: zonal opportunities are supplied and
zonal populations demand at the weights assigned to these zones, with
minimal decay adjustment, reflected by the proportional allocation
factors \(\kappa_{ij}^D\).

\begin{table}

\caption{\label{tbl-adjusted-small-system-land-use-doubly-constrained-case-allfs-access-values-for-zone2}Doubly
constrained accessible opportunities at Zone 2 for all travel decay
groups in the modified simple system.}

\centering{

\fontsize{7.5pt}{9.0pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}>{\raggedright\arraybackslash}p{\dimexpr 22.50pt -2\tabcolsep-1.5\arrayrulewidth}|>{\centering\arraybackslash}p{\dimexpr 67.50pt -2\tabcolsep-1.5\arrayrulewidth}>{\centering\arraybackslash}p{\dimexpr 67.50pt -2\tabcolsep-1.5\arrayrulewidth}>{\centering\arraybackslash}p{\dimexpr 67.50pt -2\tabcolsep-1.5\arrayrulewidth}>{\centering\arraybackslash}p{\dimexpr 67.50pt -2\tabcolsep-1.5\arrayrulewidth}>{\centering\arraybackslash}p{\dimexpr 67.50pt -2\tabcolsep-1.5\arrayrulewidth}}
\toprule
 &  &  & \multicolumn{3}{>{\centering\arraybackslash}m{\dimexpr 202.50pt -2\tabcolsep-1.5\arrayrulewidth}}{V\textsubscript{\{ij\}}\textsuperscript{D}} \\ 
\cmidrule(lr){4-6}
 &  &  & f\textsubscript{1} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{3} & f\textsubscript{2} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{2} & f\textsubscript{3} (c\textsubscript{ij}) = 1/c\textsubscript{ij}\textsuperscript{0.1} \\ 
\cmidrule(lr){4-4} \cmidrule(lr){5-5} \cmidrule(lr){6-6}
Dest. & Population at 2 (units: \emph{people in 10,000s}) & Opportunities (units: \emph{capacity in 10,000s}) & units: \emph{physician-capacity in 10,000s} & units: \emph{physician-capacity in 10,000s} & units: \emph{physician-capacity in 10,000s} \\ 
\midrule\addlinespace[2.5pt]
1 & 10.000 & 7.000 & 2.133 & 2.272 & 3.411 \\ 
2 & 10.000 & 5.000 & 4.959 & 4.766 & 2.667 \\ 
3 & 10.000 & 8.000 & 2.904 & 2.958 & 3.919 \\ 
\bottomrule
\end{tabular*}

}

\end{table}%

Recall, accessibility is defined as ``the potential for interaction''
and is traditionally presented as a summary zonal measure. In the doubly
constrained case, we force the zonal population and zonal opportunities
to match one-to-one, hence providing a zonal summary is no longer
relevant: the sum of \(V_{ij}^D\) for all \(i\)s ends up being equal to
the population at the zone. However, if what interests readers is the
``potential for interaction'' in the case population and opportunities
to match one-to-one, reframing the investigation may be needed. In this
sense, it would be examining ``interaction'' (much less room for
potential) through the values of \(V_{ij}^D\) e.g., how many
opportunities are being allocated to an origin from a destination (or in
a transposed sense for \(M_{ji}^D\)). In this sense, the doubly
constrained case can be thought of as an estimate of \emph{realized}
accessibility or access: reflecting spatial interaction. It is also
formulaically identical to the doubly constrained spatial interaction
model, but with specific interpretations of the origin and destination
weights as `population' and `opportunity', respectively.

As Wilson explicitly noted, origin and destination weights defined in
the spatial interaction model \emph{can} be defined using any unit.
Accessibility, however, is often presented and understood as a zonal
summary of \emph{potential} for interaction between origins and
destinations that contain inherently different units. Arriving at the
doubly constrained case through the unconstrained, total, and singly
constrained cases make the connection between the \emph{potential} for
spatial interaction (accessibility) and \emph{realized} potential for
spatial interaction more interpretable. Namely, by demonstrating that
these members of the accessibility family all derive from the same root
and can be derived from Wilson's original formulation, this more clearly
demonstrates what \emph{potential} is, within the context of spatial
interaction modelling.

Namely, \emph{potential} depends on the framing of the masses at the
origin and destination, and how similar they are in their units. The
appropriate constraint should be decided based on the the input data and
their similarity. In increasing unit similarity from the perspective of
opportunity-constraint: the total-constraint can be used if population
(origin mass) is not known, the single constraint can be used if
population is known and matters to \emph{potential} but it does not
match the capacity of opportunities, and lastly, the double constraint
is appropriate if population is known, matters and population matches
the capacity of opportunities. These constraint measures can reflect the
opportunities accessible at each zone, but reflects the assumptions
embedded in the input data about the potential to spatially interact.

\section{Conclusions}\label{conclusions}

In this work, we examined the historical and mathematical commonalities
between spatial interaction models and place-based accessibility
measures. As accessibility research evolved largely influenced by Hansen
\citep{hansen1959}, researchers in the field neglected the
proportionality constant that was originally present in gravity-based
models, and is still present in spatial interaction modeling. Firstly,
we have demonstrated that by reintroducing such constraints and defining
associated balancing factors and proportional allocation factors, we can
derive a unifying family of accessibility measures that reintroduces
units to the resulting values. These values become more intuitive for
the purpose of analysis and comparison. Secondly, the constraining
constants, depending on the case (i.e., total, singly- or doubly-
constrained), restrict the degree of \emph{potential}, linking
accessibility (the potential for spatial interaction) with access
(spatial interaction) on the same continuum based on the constraint
used. Lastly, we discussed how popular measures such as the one used in
Hansen \citep{hansen1959} and the 2SFCA link into the family of
accessibility measures. We summarise the family of accessibility
measures, as follows.

We first place the popular Hansen-type accessibility measure
\citep{hansen1959} within this family of measures as an
``unconstrained'' case, demonstrating that resulting values cannot be
directly compared across different travel scenarios without ad-hoc
adjustments. We then show how applying a total constraint balances the
units and produces a statistically averaged solution that converges to
the regional average for each zone as the decay effect decreases. In
other words, the total-constraint model could be a more interpretable
alternative for the unconstrained case if population-competition is not
relevant and one is interested in capturing the maximum
\emph{potential}; specifically, if there is a fixed number of
opportunities in the region, and if it makes sense to assume that people
accessing proximate opportunities leave fewer for others, \emph{without}
considering the population size at the origins.

We then introduce the singly constrained case, which \emph{does} takes
into account the population size at the origin in the allocation of
opportunities (unlike the total constraint). It is also mathematically
equivalent to the spatial availability introduced in Soukhov et al.
\citep{soukhovIntroducingSpatialAvailability2023}. In this case, all
accessibility values are fixed to sum to a known zonal opportunity-size
value (implicitly, the regional total of opportunities), but they are
not required to sum to any population-based values at the zone or
regional level. The singly constrained model could be useful if regional
competition is a factor and if the acknowledgment that only a finite
number of opportunities can be allocated from each destination (with
those allocations distributed based on origin population size) is
suitable. We also introduce an `accessible' PPR (e.g., opportunities per
capita), calculated by dividing each accessibility value by the zonal
population. To clarify, this per capita expression of the singly
constrained case is equivalent to the 2SFCA \citep{luo2003, shen1998},
hence linking this literature back to spatial interaction principles.

Lastly, the doubly constrained case is introduced. In this case, the
sums must equal both the regional total and ensure that no zone
allocates more opportunities than it has available. Specifically,
accessibility values for each \(i\)-\(j\) pair must be a proportion of
he zonal opportunity and population values simultaneously. For example,
the accessibility at zone 1 must equal the sum of opportunities from
zones 1, 2, and 3, as well as the sum of the population at zone 1.
Satisfying the double constraint means the opportunities and population
data must match one-to-one, so working with the accessibility
\(i\)-\(j\) pair values should be of interest. In this sense, the
research question should be concerned with `access' (how many
opportunities accessed from \(j\) at \(i\) based on given zonal
opportunities and populations) instead of potential spatial interaction
(e.g., typically expressed as a zonal summary measure of how many
opportunities one could reach (out of a regional total and/or
zonal-allocation).

In summary, building on Wilson's \citep{wilson1971} foundational work,
this paper proposed a unified framework for accessibility that is able
to account for competition. By reintroducing Wilson's proportionality
constant, the proposed family of constrained accessibility measures
restores measurement units to accessibility estimates. This enhancement
provides a more interpretable, consistent, and theoretically grounded
basis for accessibility analysis, which could help advance the adoption
of accessibility-oriented planning.

\nolinenumbers
\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}